---
title: "Functional annotations"
author: "Miloš Duchoslav"
date: "2025-11"
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Setting NO evaluation of code chunks as default (needed for knitting of bash code without trying to run that)
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction

This RMarkdown file (or its markdown version for GitHub) documents compiling functional annotation from several sources.

This file includes:

1. BASH code that I ran at MetaCentrum (Czech national grid infrastructure) running PBS scheduling system for batch jobs.
2. R code that I ran locally.

### SW installation and versions

The SW installation instructions and versions of SW used is described in [Installation_of_SW.md](Installation_of_SW.md).

# Strategy

Sources of functional annotation:

1. TAIR ([arabidopsis.org](https://www.arabidopsis.org/)) functional annotations for *A. thaliana* homologues
2. [plantcyc.org](https://plantcyc.org/) functional annotations for *A. thaliana* homologues
3. [UniProt](https://www.uniprot.org/uniprot/?query=proteome%3Aup000006548) functional annotations for *A. thaliana* homologues
4. [InterProScan](https://interproscan-docs.readthedocs.io) results for protein sequences of the given species

In case of several *A. thaliana* homologues assigned to one gene, annotation of all of them was compiled together.

## Detailed sources of *A. thaliana* annotations

Extracted from script [functional_annotation_compilation_executable.r](functional_annotation_compilation_executable.r).

```
## Gene aliases
# File "gene_aliases_20241001.txt.gz"
# Downloaded on 2024-12-09 from
# https://v2.arabidopsis.org/download_files/Subscriber_Data_Releases/TAIR_Data_20240930/gene_aliases_20241001.txt.gz
# (web page https://v2.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FSubscriber_Data_Releases%2FTAIR_Data_20240930)

## Subcellular predictions
# File "Araport11-Subcellular_Predictions_version_2024_03_09.txt"
# Downloaded on 2024-12-09 from https://www.arabidopsis.org/download/list?dir=Genes%2FAraport11_genome_release

## Functional descriptions from arabidopsis.org
# File "Araport11_functional_descriptions_20241001.txt.gz"
# Downloaded on 2024-12-09 from
# https://v2.arabidopsis.org/download_files/Subscriber_Data_Releases/TAIR_Data_20240930/Araport11_functional_descriptions_20241001.txt.gz 
# (web page https://v2.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FSubscriber_Data_Releases%2FTAIR_Data_20240930)

## GO annotations from arabidopsis.org
# Downloaded on 2025-11-11 from
# https://www.arabidopsis.org/download/file?path=Public_Data_Releases/TAIR_Data_20240930/ATH_GO_GOSLIM.txt.gz 
# (web page https://www.arabidopsis.org/download/list?dir=Public_Data_Releases%2FTAIR_Data_20240930)

## Metabolic pathways
# File "aracyc_pathways.20230103"
# Downloaded on 2024-12-09 from
# https://plantcyc-ftp.storage.googleapis.com/pmn/Pathways/Data_dumps/PMN15.5_January2023/pathways/aracyc_pathways.20230103
# (web page https://plantcyc.org/)

## Annotations from Uniprot
# File "uniprotkb_proteome_up000006548_2025_11_11.tsv.gz"
# Downloaded on 2025-11-11 from https://www.uniprot.org/uniprot/?query=proteome%3Aup000006548.
# I selected the desired columns for view and then downloaded that as a table.
```


# InterProScan - direct annotation of the genes of desired species (not *A. thaliana* orthologues)

Inspiration from:

Lian, Qichao, Bruno Huettel, Birgit Walkemeier, Baptiste Mayjonade, Céline Lopez-Roques, Lisa Gil, Fabrice Roux, Korbinian Schneeberger, and Raphael Mercier. “A Pan-Genome of 69 Arabidopsis Thaliana Accessions Reveals a Conserved Genome Structure throughout the Global Species Range.” Nature Genetics 56, no. 5 (May 2024): 982–91. https://doi.org/10.1038/s41588-024-01715-9.:

The resulting gene models were further annotated functionally using InterProScan v5.59-91.0 (ref. 109) (parameters: -f TSV -t p -iprlookup -goterms -pa).



## InterProScan - preparation of input files

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/
mkdir interproscan

cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

## Removing asterisks from protein sequences

# Interproscan doesn't like * in protein sequences, even in the end. We have to remove them from theprotein sequences.

# checking for internal stop codons (* in the middle of sequence)
for file in ../primary_transcripts/*.fasta
do
STOPS=$(awk '/^>/ { if(NR>1) print "";  printf("%s\n",$0); next; } { printf("%s",$0);}  END {printf("\n");}' < $file | grep \*[[:alpha:]] | wc -l)
echo $file $STOPS
done

# checking for * anywhere
grep '\*' ../primary_transcripts/Arabidopsis_arenosa.fasta | wc -l # 33901

# check of sed command
sed -E 's/\*//g' ../primary_transcripts/Noccaea_praecox.fasta | less

## remove all asterisks from protein sequences
# Note: This will remove also internal stop codons in bad proteins.

mkdir -p primary_transcripts_without_stops

for file in ../primary_transcripts/*.fasta
do
  # get filename without path
  filename=$(echo $file | sed 's,^.*/,,')
  
  # remove all asterisks from lines that do not start with >
  sed -E 's/\*//g' ../primary_transcripts/$filename > primary_transcripts_without_stops/$filename
done


```

## InterProScan run

```{sh}
### Metacentrum script

#PBS -N InterProScan_all_species
#PBS -l select=1:ncpus=8:mem=96gb:scratch_local=200gb
#PBS -l walltime=24:00:00 
#PBS -m ae

## It is needed to set the $species variable during submitting the job like this:
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Cardamine_glauca Noccaea_praecox
# do
# echo "Submitting job for species: $species"
# qsub  -v "species=$species" InterProScan_all_species.bash
# done

echo "Species for this job: $species"

# define variables
query_fasta="/storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan/primary_transcripts_without_stops/$species.fasta"
output_dir=/storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

# append a line to a file "jobs_info.txt" containing the ID of the job, the hostname of node it is run on and the path to a scratch directory
# this information helps to find a scratch directory in case the job fails and you need to remove the scratch directory manually 
echo "$PBS_JOBID is running on node `hostname -f` in a scratch directory $SCRATCHDIR" | ts '[%Y-%m-%d %H:%M:%S]' >> $PBS_O_WORKDIR/jobs_info.txt

# test if scratch directory is set
# if scratch directory is not set, issue error message and exit
test -n "$SCRATCHDIR" || { echo >&2 "Variable SCRATCHDIR is not set!"; exit 1; }

# move into scratch directory
cd $SCRATCHDIR 

# load Java
module load openjdk
module list

# run InterProScan
/storage/brno12-cerit/home/duchmil/SW/InterProScan/interproscan-5.76-107.0/interproscan.sh --goterms --pathways --enable-tsv-residue-annot --tempdir $SCRATCHDIR --cpu 8 -i $query_fasta --output-file-base $species

echo "Main calculation done." | ts '[%Y-%m-%d %H:%M:%S]'

# compress output files
pigz -v -p 8 $species.*

# move the output to user's DATADIR or exit in case of failure
mkdir -p $output_dir # will make folder if it doesn't exist
cp -v $species.* $output_dir/ || { echo >&2 "Result file(s) copying failed (with a code $?) !!"; exit 4; }

echo "Copying output file to $output_dir done." | ts '[%Y-%m-%d %H:%M:%S]'

# clean the SCRATCH directory
clean_scratch

# Resources: From 2 h (A. thaliana, probably most precomputed results) up to 24 h (Cochlearia excelsa, many proteins). For most around 10 h, 60-90 % CPU, 50-96 GB memory.

```

### Loop for submitting jobs for several species
```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/metacentrum_scripts

# for species in Aethionema_saxatile Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Arabidopsis_thaliana Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_amara Cardamine_glauca Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Erysimum_linariifolium Euclidium_syriacum Eutrema_salsugineum Noccaea_praecox Odontarrhena_muralis Raphanus_sativus
# for species in Aethionema_saxatile
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Cardamine_amara Cardamine_glauca Erysimum_linariifolium Noccaea_praecox Odontarrhena_muralis
for species in Arabidopsis_thaliana Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Euclidium_syriacum Eutrema_salsugineum Raphanus_sativus
do
echo "Submitting job for species: $species"
qsub  -v "species=$species" InterProScan_all_species.bash
done


```

### Interproscan results

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

# view beginning as table
zcat Arabidopsis_arenosa.tsv.gz | head -n 1000 | column -t -s $'\t' |  less -S

zcat Arabidopsis_arenosa.tsv.gz | wc -l # 315161 (older run with older version of Interproscan: 313362)

zcat Arabidopsis_arenosa.tsv.gz | cut -f 1 | sort | uniq | wc -l # 31121 (out of 33901 proteins) (older run with older version of Interproscan: 31111)

# sites
zless -S Arabidopsis_arenosa.tsv.sites.gz

zcat Arabidopsis_arenosa.tsv.sites.gz | cut -f 1 | sort | uniq | wc -l # 11409 (out of 33901 proteins)
```
#### TSV description
https://interproscan-docs.readthedocs.io/en/latest/OutputFormats.html

The TSV format presents the match data in columns as follows:

    Protein accession (e.g. P51587)

    Sequence MD5 digest (e.g. 14086411a2cdf1c4cba63020e1622579)

    Sequence length (e.g. 3418)

    Analysis (e.g. Pfam / PRINTS / Gene3D)

    Signature accession (e.g. PF09103 / G3DSA:2.40.50.140)

    Signature description (e.g. BRCA2 repeat profile)

    Start location

    Stop location

    Score - is the e-value (or score) of the match reported by member database method (e.g. 3.1E-52)

    Status - is the status of the match (T: true)

    Date - is the date of the run

    InterPro annotations - accession (e.g. IPR002093)

    InterPro annotations - description (e.g. BRCA2 repeat)

    GO annotations with their source(s), e.g. GO:0005515(InterPro)|GO:0006302(PANTHER)|GO:0007195(InterPro,PANTHER). This is an optional column; only displayed if the --goterms option is switched on

    Pathways annotations, e.g. REACT_71. This is an optional column; only displayed if the --pathways option is switched on

#### Analysis description

                      TIGRFAM (XX.X) : TIGRFAMs are protein families based on hidden Markov models (HMMs).
                         SFLD (X) : SFLD is a database of protein families based on hidden Markov models (HMMs).
                  SUPERFAMILY (X.XX) : SUPERFAMILY is a database of structural and functional annotations for all proteins and genomes.
                      PANTHER (XX.X) : The PANTHER (Protein ANalysis THrough Evolutionary Relationships) Classification System is a unique resource that classifies genes by their functions, using published scientific experimental evidence and evolutionary relationships to predict function even in the absence of direct experimental evidence.
                       Gene3D (X.X.X) : Structural assignment for whole genes and genomes using the CATH domain structure database.
                        Hamap (XXXX_XX) : High-quality Automated and Manual Annotation of Microbial Proteomes.
                  ProSiteProfiles (XXX_XX) : PROSITE consists of documentation entries describing protein domains, families and functional sites as well as associated patterns and profiles to identify them.
                        Coils (X.X.X) : Prediction of coiled coil regions in proteins.
                        SMART (X.X) : SMART allows the identification and analysis of domain architectures based on hidden Markov models (HMMs).
                          CDD (X.XX) : CDD predicts protein domains and families based on a collection of well-annotated multiple sequence alignment models.
                       PRINTS (XX.X) : A compendium of protein fingerprints - a fingerprint is a group of conserved motifs used to characterise a protein family.
                        PIRSR (XXXX_XX) : PIRSR is a database of protein families based on hidden Markov models (HMMs) and Site Rules.
                  ProSitePatterns (XXXX_XX) : PROSITE consists of documentation entries describing protein domains, families and functional sites as well as associated patterns and profiles to identify them.
                      AntiFam (X.X) : AntiFam is a resource of profile-HMMs designed to identify spurious protein predictions.
                         Pfam (XX.X) : A large collection of protein families, each represented by multiple sequence alignments and hidden Markov models (HMMs).
                   MobiDBLite (X.X) : Prediction of intrinsically disordered regions in proteins.
                        PIRSF (X.XX) : The PIRSF concept is used as a guiding principle to provide comprehensive and non-overlapping clustering of UniProtKB sequences into a hierarchical order to reflect their evolutionary relationships.
  

# Functional annotations of *Arabidopsis thaliana* orthologs


## Compilation of annotations from several sources

The compilation is done by executable R script [functional_annotation_compilation_executable.r](functional_annotation_compilation_executable.r). The script is not optimised for speed, the computation can take up to 2 h per species. Extraction of the commands to the executable script allowed me to run it distantly in parallel jobs.

The script includes information about data sources.

### List of species

```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
old.par<-par(no.readonly = T)

# read species names
interpro.files <- list.files(path = "interproscan/", pattern = ".tsv.gz")
species <- sub(pattern = ".tsv.gz", replacement = "", x = interpro.files, fixed = T)
species
```

### Testing of script

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3

# load R
module load r/4.1.3-gcc-10.2.1-6xt26dl

Rscript --verbose "functional_annotation_compilation_executable.r" Aethionema_saxatile
```

### Bash script for running of R script (compilation of functional annotations)

```{sh}
### Metacentrum script

#PBS -N functional_annotation_compilation
#PBS -l select=1:ncpus=1:mem=32gb:scratch_local=200gb
#PBS -l walltime=24:00:00 
#PBS -m ae

## It is needed to set the $species variable during submitting the job like this:
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Cardamine_glauca Noccaea_praecox
# do
# echo "Submitting job for species: $species"
# qsub  -v "species=$species" InterProScan_all_species.bash
# done

echo "Species for this job: $species" | ts '[%Y-%m-%d %H:%M:%S]'

# append a line to a file "jobs_info.txt" containing the ID of the job, the hostname of node it is run on and the path to a scratch directory
# this information helps to find a scratch directory in case the job fails and you need to remove the scratch directory manually 
echo "$PBS_JOBID is running on node `hostname -f` in a scratch directory $SCRATCHDIR" | ts '[%Y-%m-%d %H:%M:%S]' >> $PBS_O_WORKDIR/jobs_info.txt

# move to working dir
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3

# load R
module load r/4.1.3-gcc-10.2.1-6xt26dl

# run R script
Rscript --verbose "functional_annotation_compilation_executable.r" $species || { echo >&2 "Rscript failed!"; exit 1; }

echo "Run ended: $species" | ts '[%Y-%m-%d %H:%M:%S]'

# clean the SCRATCH directory
clean_scratch

# Resources: 100 % CPU, up to 2 h, 6 GB memory.
```

### Loop for submitting jobs for several species

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/metacentrum_scripts

# for species in Aethionema_saxatile Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_amara Cardamine_glauca Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Erysimum_linariifolium Euclidium_syriacum Eutrema_salsugineum Noccaea_praecox Odontarrhena_muralis Raphanus_sativus
# for species in Alyssum_gmelinii Arabidopsis_arenosa
for species in Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_amara Cardamine_glauca Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Erysimum_linariifolium Euclidium_syriacum Eutrema_salsugineum Noccaea_praecox Odontarrhena_muralis Raphanus_sativus
do
echo "Submitting job for species: $species"
qsub  -v "species=$species" functional_annotation_compilation.bash
done


```


# Adding reliability table for species annotated by me

The script at the same time plots histograms of protein length for different categories of protein reliability.

```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
if(!exists("old.par")) old.par<-par(no.readonly = T)

# dirs with annotations
dirs.list <- list.dirs(path = "D:/!ecolgen/annotations/", recursive = F)
# support tables
files.list <- list.files(path = paste0(dirs.list, "/final_files/"), pattern = "_protein_coding_genes_support.tsv", full.names = T)

## preparation of dataframe for histogram data
# set boundaries and coarsness of frequency calculation
l.min <- 0
l.max <- 10000
l.by <- 25
# prepare data frames
length.hist <- data.frame(bins = seq(from = l.min + (l.by / 2), to = l.max - (l.by / 2), by = l.by))
length.hist.1 <- data.frame(bins = seq(from = l.min + (l.by / 2), to = l.max - (l.by / 2), by = l.by))
length.hist.2 <- data.frame(bins = seq(from = l.min + (l.by / 2), to = l.max - (l.by / 2), by = l.by))

# write files?
write.files <- F

## Loop through the species
# Add reliability table and collect data for histograms

for(i in seq_along(files.list)) {
	
	supp <- read.table(file = files.list[i], header = T, sep = "\t")
	# supp <- read.table(file = "final_files/Odontarrhena_muralis_CUNI_V1_annotation_v1.0_protein_coding_genes_support.tsv", header = T, sep = "\t")
	
	summary(supp)
	
	colnames(supp) <- paste0("support_", colnames(supp))
	
	# Extract species from file name
	file.name <- gsub(pattern = "^.*/", replacement = "", x = files.list[i])
	one.species <- gsub(pattern = "(^[^_]*_[^_]*).*", replacement = "\\1", x = file.name)
	
	# read RDS
	annotations.3 <- readRDS(file = paste0("functional_annotation/4_full_rds/", one.species, "_At_orthologues_and_functional_annotations_full.rds"))
	
	colnames(annotations.3)
	
	# merge tables
	annotations.4 <- merge(x = annotations.3, y = supp, by.x = one.species, by.y = "support_Gene", all = T)
	
	# column after which to move the supp columns
	d <- 2
	# move supp columns
	annotations.5 <- annotations.4[, c(1:d, (ncol(annotations.3) + 1):ncol(annotations.4), (d + 1):ncol(annotations.3))]
	
	colnames(annotations.5)
	
	# save merged table
	if(write.files) {
		# make folder
		dir.create(path = "functional_annotation/5_full_with_support_for_my_annotations_rds", recursive = T, showWarnings = F)
		# write RDS
		saveRDS(object = annotations.5, 
						file = paste0("functional_annotation/5_full_with_support_for_my_annotations_rds/", one.species, "_support_and_At_orthologues_and_functional_annotations.rds"))
	}
	
	# Genes supported by proteins from at least one species
	support_any_protein <- rowSums(annotations.5[, c("support_A.lyrata.pep", "support_A.thaliana.pep", "support_B.rapa.pep")]) >= 1

	# Genes with any support (either protein from at least one species or transcripts assembled from RNAseq)
	support_any <- rowSums(annotations.5[, c("support_Assembled_transcripts", "support_A.lyrata.pep", "support_A.thaliana.pep", "support_B.rapa.pep")]) >= 1

	# Genes with support of proteins from all species and transcripts
	support_all <- rowSums(annotations.5[, c("support_Assembled_transcripts", "support_A.lyrata.pep", "support_A.thaliana.pep", "support_B.rapa.pep")]) == 4
	
	# calculate data for histograms
	# hist(annotations.5$prot_length[annotations.5$support_Assembled_transcripts == 1], breaks = 50)
	# hist(annotations.5$prot_length[support_any == 1], breaks = 50)
	
	# all proteins
	spec.hist <- hist(annotations.5$prot_length, breaks = seq(from = l.min, to = l.max, by = l.by), xlim = c(0, 3000), main = one.species)
	# proteins with any support (either protein from at least one species or transcripts assembled from RNAseq)
	spec.hist.1 <- hist(annotations.5$prot_length[support_any == 1], breaks = seq(from = l.min, to = l.max, by = l.by), xlim = c(0, 3000), main = one.species)
	# proteins with support of proteins from all species and transcripts
	spec.hist.2 <- hist(annotations.5$prot_length[support_all == 1], breaks = seq(from = l.min, to = l.max, by = l.by), xlim = c(0, 3000), main = one.species)
	
	# save to prepared dataframe
	length.hist[, one.species] <- spec.hist$density
	length.hist.1[, one.species] <- spec.hist.1$density
	length.hist.2[, one.species] <- spec.hist.2$density
	
}

## Plotting histograms
pdf ("figs_and_stats/prot_length_histogram_supported_proteins.pdf", width=7, height=7, onefile = T)

# plots input
g.input <- c("length.hist", "length.hist.1", "length.hist.2")
g.sub <- c("All genes", 
						"Genes supported by aligned protein from at least one species or by transcript",
						"Genes supported by aligned proteins from all three species and by transcript")
# all species
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19, lwd = 2,
					main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length (amino acid residues)")
	legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1, lwd = 2)
	mtext(text = g.sub[i])
}

# Resource paper
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19,
				main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length (amino acid residues)", lwd = 2)
	legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1, lwd = 2)
	mtext(text = g.sub[i])
}
dev.off()

## Plot for publication
# use palette from ColorBrewer
require(RColorBrewer)
display.brewer.all(n = 5, type = "qual", colorblindFriendly = T)

# plots input
g.input <- c("length.hist", "length.hist.1", "length.hist.2")
g.sub <- c("All genes", 
						"Genes supported by aligned protein from at least one species or by transcript",
						"Genes supported by aligned proteins from all three species and by transcript")

# Resource paper
pdf ("figs_and_stats/prot_length_histogram_for_publication_supported_proteins.pdf", width=7, height=7, onefile = T)
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3))
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	mtext(text = g.sub[i], line = 2)
}
dev.off()

pdf ("figs_and_stats/prot_length_histogram_for_publication_supported_proteins_2.pdf", width=10, height=5, onefile = T)
par(mfrow = c(1, 2))
for(i in 1:2) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3), las = 1)
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	# mtext(text = g.sub[i], line = 2)
	mtext(text = c("A", "B")[i], line = 2, adj = 0, font = 2, cex = 2)
}
par(old.par)
dev.off()

svg ("figs_and_stats/prot_length_histogram_for_publication_supported_proteins_3.svg", width=180/25.4, height=90/25.4, onefile = T)
par(mfrow = c(1, 2), cex = 0.75)
for(i in 1:2) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3), las = 1)
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	# mtext(text = g.sub[i], line = 2)
		# mtext(text = c("a", "b")[i], line = 2, adj = 0, font = 2, cex = 2)
	mtext(text = c(expression(bold("a") * " All gene models"), expression(bold("b") * " Gene models with external evidence"))[i], line = 1.5, adj = 0, font = 1, cex = 1, outer = F)
}
par(old.par)
dev.off()

# just range of short proteins
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 500), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3))
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	mtext(text = g.sub[i], line = 2)
}

```


# Description of output tables

There are several versions of tables (with different subsets of columns) in the [functional_annotation directory](functional_annotation/) and you can use the one which suits you:

1. [1_only_At_orthologues_and_GO](functional_annotation/1_only_At_orthologues_and_GO/)
	- Only the main columns: `Species_name` (gene IDs of the given species), `prot_length`, `Arabidopsis_thaliana`, `single_Arabidopsis_thaliana`, `homologue_type`, `GO_term_IDs` and `UniProt_Protein.names`
	- Format: plain tsv
2. [2_full_without_InterProScan_pathways_tsv_gz](functional_annotation/2_full_without_InterProScan_pathways_tsv_gz/)
	- All columns, just without `InterProScan_pathways` column, which contains a lot of data, making files too big.
	- Format: gzipped tsv
3. [3_InterProScan_pathways_tsv_gz](functional_annotation/3_InterProScan_pathways_tsv_gz/)
	- Only `Species_name` (gene IDs of the given species) and `InterProScan_pathways`
	- Format: gzipped tsv
4. [4_full_rds](functional_annotation/4_full_rds/)
	- All columns
	- Format: rds (can be imported to R using function `readRDS()`)
5. [5_full_with_support_for_my_annotations_rds](functional_annotation/5_full_with_support_for_my_annotations_rds/)
	- All columns plus gene model support columns (only for genomes annotated by me)
	- Format: rds (can be imported to R using function `readRDS()`)
	
## Explanation of columns in the output tables

Columns:

- **Main columns**  
	- `Species_name` (e.g. `Arabidopsis_arenosa`): Gene IDs of the investigated species.  
	- `prot_length`: Length of protein sequence of the longest/primary isoform used for OrthoFinder.  
	- `Arabidopsis_thaliana`: Selected *A. thaliana* orthologues according to these rules:  
				1. Use orthologues from orthologues table.  
				2. If missing, use Blast RBH.  
				3. If missing, use genes from N0 hierarchical orthogroups.  
				4. If missing, use genes from broad orthogroups.  
				5. If missing, use genes hits from simple Blast (but only those with BLAST_pident > 40 & BLAST_qcovhsp > 50)  
						BLAST_pident: Percentage of identical matches (in local alignment)  
						BLAST_qcovhsp: Query coverage per HSP (%)  
								HSP = High-scoring Segment Pair (local alignment with no gaps)  
	- `single_Arabidopsis_thaliana`: Selected *A. thaliana* orthologue (only one) according to these rules:  
				1. If there is only one *A. thaliana* gene in column `Arabidopsis_thaliana`, use that one.  
				2. Else, if there is one and only RBH and it is among genes in column `Arabidopsis_thaliana`, take the RBH.  
				3. Else, take the gene from the column `Arabidopsis_thaliana` that had the highest bitscore in the simple BLAST results. If several have the highest bitscore, take the first.  
				4. If the genes in column `Arabidopsis_thaliana` are not among BLAST hits, take nothing.  
					`Recommendation`: It is appealing to use the "single" orthologues (`single_Arabidopsis_thaliana`), because it makes things easier. However, I recommend to do that only in cases when it is absolutely necessary and otherwise use the column where there are multiple orthologues in some cases (`Arabidopsis_thaliana`). The real orthology is not always one-to-one due to multiplications of genes in some species (there is a good explanation in [OrthoFinder GitHub](https://github.com/davidemms/OrthoFinder)). The column `Arabidopsis_thaliana` should better reflect the real biology.
	- `homologue_type`: Type of selected *A. thaliana* orthologues in `Arabidopsis_thaliana` column.  
			- *orthologue* - orthologue from OrthoFinder orthologues table (method 1)  
			- *other* - homologue assigned by other method (2-5)  
	- `GO_term_IDs`: Gene ontology term IDs compiled from columns `TAIR_GO`, `UniProt_Gene.Ontology.IDs` and `InterProScan_GO_term`.  
- **OrthoFinder_[...]**: Results from OrthoFinder.  
	- `OrthoFinder_OL_Arabidopsis_thaliana` - *A. thaliana* orthologues from OrthoFinder orthologues table.  
	- `OrthoFinder_N0_HOG` - N0 orthogroup ID from OrthoFinder N0 hierarchical orthogroups table.  
	- `OrthoFinder_N0_Arabidopsis_thaliana` - *A. thaliana* orthologues/homologues from OrthoFinder N0 hierarchical orthogroups table.  
	- `OrthoFinder_OG_Orthogroup` - Orthogroup ID from OrthoFinder orthogroups table.  
	- `OrthoFinder_OG_Arabidopsis_thaliana` - *A. thaliana* orthologues/homologues from OrthoFinder orthogroups table.  
- **RBH_[...]**: Results of Reciprocal best hit (RBH) BLAST.    
	- `RBH_B_id` - ID from species B (*A. thaliana*)  
	- `RBH_A_length` - Length of protein sequence of the investigated (A) species (unfortunately including `*` in the end)  
	- `RBH_B_length` - Length of *A. thaliana* (species B) hit protein sequence (unfortunately including `*` in the end)  
	- `RBH_A_qcovhsp` - Percentage of sequence A covered by alignment (investigated species)  
	- `RBH_B_qcovhsp` - Percentage of sequence B covered by alignment (*A. thaliana*)  
	- `RBH_length` - HSP alignment length  
	- `RBH_pident` - HSP percentage identity  
	- `RBH_bitscore` - HSP bitscore  
- **BLAST_[...]**: Results for highest-scoring hit of simple BLAST (investigated species against *A. thaliana* protein data  
	- `BLAST_qaccver` - Query (investigated species) accession and version  
	- `BLAST_saccver` - Subject (*A. thaliana*) accession and version  
	- `BLAST_pident` - Percentage of identical matches  
	- `BLAST_length` - Alignment length  
	- `BLAST_mismatch` - Number of mismatches  
	- `BLAST_gapopen` - Number of gap openings  
	- `BLAST_qstart` - Start of alignment in query  
	- `BLAST_qend` - End of alignment in query  
	- `BLAST_sstart` - Start of alignment in subject (database hit)  
	- `BLAST_send` - End of alignment in subject (database hit)  
	- `BLAST_evalue` - Expectation value (E-value)  
	- `BLAST_bitscore` - Bit score  
	- `BLAST_qcovhsp` - Query coverage per HSP (alignment) [%]  
	- `BLAST_qlen` - Query sequence length  
	- `BLAST_slen` - Subject sequence length
- **TAIR_[...]**: Functional annotations obtained from TAIR ([arabidopsis.org](https://www.arabidopsis.org/)) compiled for *A. thaliana* homologues in `Arabidopsis_thaliana` column.
	- `TAIR_symbol`  
	- `TAIR_full_name`  
	- `TAIR_subcellular.prediction`  
	- `TAIR_gene_model_type`  
	- `TAIR_short_description`  
	- `TAIR_Curator_summary`  
	- `TAIR_Computational_description`  
	- `TAIR_GO`
- **PlantCyc_pathway**: Metabolic pathways obtained from [plantcyc.org](https://plantcyc.org/) compiled for *A. thaliana* homologues in `Arabidopsis_thaliana` column.
	- `PlantCyc_pathway`  
- **TAIR_[...]**: Annotations obtained from [UniProt](https://www.uniprot.org/uniprot/?query=proteome%3Aup000006548) compiled for *A. thaliana* homologues in `Arabidopsis_thaliana` column.
	- `UniProt_Entry`  
	- `UniProt_Entry.Name`  
	- `UniProt_Protein.names`  
	- `UniProt_Gene.Names`  
	- `UniProt_Length`  
	- `UniProt_Gene.Names..ordered.locus.`  
	- `UniProt_Gene.Names..primary.`  
	- `UniProt_Gene.Names..synonym.`  
	- `UniProt_Pathway`  
	- `UniProt_Cofactor`  
	- `UniProt_DNA.binding`  
	- `UniProt_Binding.site`  
	- `UniProt_Active.site`  
	- `UniProt_Protein.existence`  
	- `UniProt_Keywords`  
	- `UniProt_Interacts.with`  
	- `UniProt_Developmental.stage`  
	- `UniProt_Induction`  
	- `UniProt_Tissue.specificity`  
	- `UniProt_Gene.Ontology..biological.process.`  
	- `UniProt_Gene.Ontology..GO.`  
	- `UniProt_Gene.Ontology.IDs`  
	- `UniProt_Gene.Ontology..molecular.function.`  
	- `UniProt_Gene.Ontology..cellular.component.`  
	- `UniProt_Intramembrane`  
	- `UniProt_Subcellular.location..CC.`  
	- `UniProt_Topological.domain`  
	- `UniProt_Transmembrane`  
	- `UniProt_PROSITE`  
	- `UniProt_Pfam`  
	- `UniProt_InterPro`  
	- `UniProt_Subunit.structure`  
	- `UniProt_thalemine.link`  
	- `UniProt_tair.link`  
- **InterProScan_[...]**: Results of InterProScan for protein sequences of the given species (NOT *A. thaliana* homologues).
	- `InterProScan_AntiFam`  
	- `InterProScan_CDD`  
	- `InterProScan_Coils`  
	- `InterProScan_FunFam`  
	- `InterProScan_Gene3D`  
	- `InterProScan_Hamap`  
	- `InterProScan_MobiDBLite`  
	- `InterProScan_NCBIfam`  
	- `InterProScan_PANTHER`  
	- `InterProScan_PIRSF`  
	- `InterProScan_PRINTS`  
	- `InterProScan_Pfam`  
	- `InterProScan_ProSitePatterns`  
	- `InterProScan_ProSiteProfiles`  
	- `InterProScan_SFLD`  
	- `InterProScan_SMART`  
	- `InterProScan_SUPERFAMILY`  
	- `InterProScan_interpro`  
	- `InterProScan_GO_term`  
	- `InterProScan_pathways`

Extra columns only for genome annotations made by me:

- **support_[...]**: Evidence for gene models obtained during genome annotation (1/0)
	- `support_Assembled_transcripts` - Overlap with transcripts assembled from RNA-seq used for annotation by StringTie (at least 90 % of gene model)  
	- `support_A.lyrata.pep` - Overlap with *Arabidopsis lyrata* proteins aligned to genome using Miniprot (the overlap should be at least 90% of both the gene model and the aligned protein)  
	- `support_A.thaliana.` - Overlap with *Arabidopsis thaliana* proteins aligned to genome using Miniprot (the overlap should be at least 90% of both the gene model and the aligned protein)  
	- `support_B.rapa.pep` - Overlap with *Brassica rapa* proteins aligned to genome using Miniprot (the overlap should be at least 90% of both the gene model and the aligned protein)  