---
title: "Statistics and plots"
author: "Miloš Duchoslav"
date: "2025-12"
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Setting NO evaluation of code chunks as default (needed for knitting of bash code without trying to run that)
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction

This RMarkdown file (or its markdown version for GitHub) documents obtaining additional statistics about inference of orthology, supplementing orthologues with other homologues and functional annotation. Some plots and statistics are already exported within the previous RMarkdown files.

### SW installation and versions

The SW installation instructions and versions of SW used is described in [Installation_of_SW.md](Installation_of_SW.md).


# Visualization of statistics from Orthofinder

## Orthogroups - species overlap


```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3")
old.par<-par(no.readonly = T)

spec.overlap <- read.table(file = "orthofinder_results/Results_brassicaceae_3/Comparative_Genomics_Statistics/Orthogroups_SpeciesOverlaps.tsv")

## heatmap with values
pdf ("figs_and_stats/Orthogroups_SpeciesOverlaps_heatmap.pdf", width=14, height=7, onefile = T)
par(mar = c(2, 12, 12, 2) + 0.1)
# input data
gdata <- spec.overlap
# turn it from similarity to distances
gdata.dist <- max(gdata) - gdata
# clustering
clust <- hclust(d = as.dist(gdata.dist), method = "single") # this method seemed best
# plot(clust)
# reordering according to clustering
gdata.clust <- gdata[clust$order, clust$order]
# reversing order
# gdata.rev <- gdata.clust[nrow(gdata.clust):1, ]
gdata.rev <- gdata.clust[, ncol(gdata.clust):1]
# image (simple heat map - heatmap has some strange coordinate system, so it is not useful)
image(t(as.matrix(gdata.rev)), xaxt = "n", yaxt = "n",
      main = "")
title(main = "Number of shared orthogroups", line = 7)
xaxs <- ((1:ncol(gdata.rev))-1)/(ncol(gdata.rev)-1)
yaxs <- ((1:nrow(gdata.rev))-1)/(nrow(gdata.rev)-1)
axis(side = 3, at = xaxs, labels = colnames(gdata.rev), las = 2, cex.axis = 0.9)
axis(side = 2, at = yaxs, labels = rownames(gdata.rev), tick = T, las = 2, cex.axis = 1)
# values
text(x = rep(x = xaxs, each = length(yaxs)), y = rep(x = yaxs, times = length(xaxs)), 
     labels = format(as.matrix(gdata.rev), big.mark = " "), 
     cex = 0.8)
mtext(text = "Diagonal: Number of orthogroups containing species.", side = 1, cex = 0.8)
par(old.par)
dev.off()



# clust <- hclust(d = as.dist(gdata.dist), method = "average")
# plot(clust)
# clust <- hclust(d = as.dist(gdata.dist), method = "ward.D")
# plot(clust)
# clust <- hclust(d = as.dist(gdata.dist), method = "ward.D2")
# plot(clust)
# 
# clust <- hclust(d = as.dist(gdata.dist), method = "complete")
# clust <- hclust(d = as.dist(gdata.dist), method = "mcquitty")
# clust <- hclust(d = as.dist(gdata.dist), method = "median")
# clust <- hclust(d = as.dist(gdata.dist), method = "centroid")
# clust$order
# plot(clust)
# colnames(gdata)[clust$order]
```

## Orthologues statistics

```{r}
folder <- "orthofinder_results/Results_brassicaceae_3/Comparative_Genomics_Statistics/"
file.pattern <- "OrthologuesStats"
files <- list.files(folder, pattern = file.pattern)

pdf ("figs_and_stats/Comparative_Genomics_Statistics_heatmaps.pdf", width=14, height=7, onefile = T)
par(mar = c(2, 12, 12, 2) + 0.1)
for(i in 5:1) {
  gdata <- read.table(file = paste0(folder, files[i]))
  # reverse rows of the matrix to be shown in the same order as the original one on picture
  gdata.rev <- gdata[nrow(gdata):1, ]
  # image (simple heat map - heatmap has some strange coordinate system, so it is not useful)
  image(t(as.matrix(gdata.rev)), xaxt = "n", yaxt = "n",
        main = "")
  title(main = sub(pattern = "\\.tsv", replacement = "", x = files[i]), line = 7)
  xaxs <- ((1:ncol(gdata.rev))-1)/(ncol(gdata.rev)-1)
  yaxs <- ((1:nrow(gdata.rev))-1)/(nrow(gdata.rev)-1)
  axis(side = 3, at = xaxs, labels = colnames(gdata.rev), las = 2, cex.axis = 0.9)
  axis(side = 2, at = yaxs, labels = rownames(gdata.rev), tick = T, las = 2, cex.axis = 1)
  # values
  text(x = rep(x = xaxs, each = length(yaxs)), y = rep(x = yaxs, times = length(xaxs)), 
       labels = format(as.matrix(gdata.rev), big.mark = " "), 
       cex = 0.8)
  mtext(text = "Note: Species in rows against species in column.", side = 1, cex = 0.8)
}
par(old.par)
dev.off()
```

## Statistics per species
```{r}
per.spec <- read.table(file = "orthofinder_results/Results_brassicaceae_3/Comparative_Genomics_Statistics/Statistics_PerSpecies.tsv", sep = "\t", quote = "", header = T, row.names = 1, nrows = 10)


## heatmap with values
pdf ("figs_and_stats/Statistics_PerSpecies_heatmap.pdf", width=16, height=7, onefile = T)
par(mar = c(2, 22, 12, 2) + 0.1)
# input data
gdata <- per.spec
# relative value per row for heatmap
gdata.rel <- gdata/rowMeans(gdata)
# rowMeans(gdata.rel)
# reversing order
gdata.rev.rel <- gdata.rel[nrow(gdata.rel):1, ]
gdata.rev <- gdata[nrow(gdata):1, ]
# gdata.rev <- gdata[, ncol(gdata):1]
# image (simple heat map - heatmap has some strange coordinate system, so it is not useful)
image(t(as.matrix(gdata.rev.rel)), xaxt = "n", yaxt = "n",
      main = "")
title(main = "Statistics per species", line = 7)
xaxs <- ((1:ncol(gdata.rev))-1)/(ncol(gdata.rev)-1)
yaxs <- ((1:nrow(gdata.rev))-1)/(nrow(gdata.rev)-1)
axis(side = 3, at = xaxs, labels = colnames(gdata.rev), las = 2, cex.axis = 0.9)
axis(side = 2, at = yaxs, labels = rownames(gdata.rev), tick = T, las = 2, cex.axis = 1)
# values
text(x = rep(x = xaxs, each = length(yaxs)), y = rep(x = yaxs, times = length(xaxs)), 
     labels = format(round(as.matrix(gdata.rev)), big.mark = " "), 
     cex = 0.8)
par(old.par)
dev.off()

```

# Histograms of protein length

Histograms of protein length for all species. For the species annotated by me, I have also histograms of protein length for different categories of protein reliability. They are documented in [3_Functional_annotations.md](3_Functional_annotations.md#adding-reliability-table-for-species-annotated-by-me).

Spectrum of protein lengths seems to nicely show an overprediction of short genes in some species. The higher-quality annotations usually don't have such a high peak in the range of short proteins. That is nicely shown by histograms of protein length for different categories of protein reliability - when we filter for higher reliability, the spectrum becomes very similar for all of the species.

```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
if(!exists("old.par")) old.par<-par(no.readonly = T)

# read the short version of annotation file
length.files <- list.files(path = "protein_length_primary_transcripts/", pattern = ".tsv")
species <- sub(pattern = "_protein_length.tsv", replacement = "", x = length.files, fixed = T)

# chosen species for testing of the loop
one.species <- "Arabidopsis_arenosa"


### Start of the  loop

# set boundaries and coarsness of frequency calculation
l.min <- 0
l.max <- 10000
l.by <- 25
# prepare data frames
length.hist <- data.frame(bins = seq(from = l.min + (l.by / 2), to = l.max - (l.by / 2), by = l.by))
# length.hist <- data.frame(bins = seq(from = 25, to = 9975, by = 50))
length.stat <- data.frame()

for(one.species in species) {
	prot.length <- read.table(file = paste0("protein_length_primary_transcripts/", one.species, "_protein_length.tsv"), header = F, sep = "\t", na.strings = "",
													comment.char = "", quote = "")
	colnames(prot.length) <- c("gene", "prot_length")

	# calculate data for histograms
	hist(prot.length$prot_length, breaks = 50)
	summary(prot.length$prot_length)
	spec.hist <- hist(prot.length$prot_length, breaks = seq(from = l.min, to = l.max, by = l.by), xlim = c(0, 3000), main = one.species)
	
	# save to prepared dataframe
	length.hist[, one.species] <- spec.hist$density

	# save statistics to prepared dataframe
	length.stat[1, "legend"] <- "Mean protein length"
	length.stat[1, one.species] <- mean(prot.length$prot_length)
	
	length.stat[2, "legend"] <- "Median protein length"
	length.stat[2, one.species] <- median(prot.length$prot_length)
}
## End of loop

## export table with statistics
dir.create(path = "figs_and_stats", recursive = T, showWarnings = F)
write.table(x = length.stat, 
						file = "figs_and_stats/protein_length_stats.tsv",
						sep = "\t", row.names = F, na = "", quote = F)

## Plotting histograms
pdf ("figs_and_stats/prot_length_histogram.pdf", width=7, height=7, onefile = T)

# all species
g.data <- length.hist
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19,
				main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Focus species
colnames(length.hist)
g.data <- length.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Arabidopsis_arenosa", "Arabidopsis_lyrata_NCBI", "Arabidopsis_lyrata_Rawat", "Arabidopsis_thaliana", "Cardamine_amara", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19,
				main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Our annotations
g.data <- length.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19,
				main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Resource paper
g.data <- length.hist[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19,
				main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length", lwd = 2)
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lwd = 2, lty = 1)

# Resource paper + well-annotated species
g.data <- length.hist[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis", "Arabidopsis_lyrata_NCBI", "Arabidopsis_thaliana", "Brassica_rapa")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19,
				main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length", lwd = 2)
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lwd = 2, lty = 1)

dev.off()

## Plot for publication
# use palette from ColorBrewer
require(RColorBrewer)
display.brewer.all(n = 5, type = "qual", colorblindFriendly = T)
# Resource paper
g.data <- length.hist[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
# Change to %
g.data[, -1] <- g.data[, -1] * 100

pdf ("figs_and_stats/prot_length_histogram_for_publication.pdf", width=7, height=7, onefile = T)
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
				main = "", ylab = "Frequency (%)", xlab = "Protein length", lwd = 2)
g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
dev.off()
```


# Statistics of supplementation of orthologues

The statistics of supplementation of orthologues is documented in [2_Supplementing_orthologues.md](2_Supplementing_orthologues.md#big-loop).


# Statistics of functional annotation

## GO terms

Counting GO terms per gene, plotting the histograms

```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
if(!exists("old.par")) old.par<-par(no.readonly = T)

# read the short version of annotation file
ann.files <- list.files(path = "functional_annotation/1_only_At_orthologues_and_GO/", pattern = ".tsv")
species <- sub(pattern = "_At_orthologues_and_GO.tsv", replacement = "", x = ann.files, fixed = T)

# chosen species for testing of the loop
one.species <- "Arabidopsis_arenosa"


### Start of the  loop

# prepare data frames
go.hist <- data.frame(bins = 0:99)
go.stat <- data.frame()

for(one.species in species) {
	short.ann <- read.table(file = paste0("functional_annotation/1_only_At_orthologues_and_GO/", one.species, "_At_orthologues_and_GO.tsv"), header = T, sep = "\t", na.strings = "",
													comment.char = "", quote = "")

	# number of GO terms per gene
	go.num <- sapply(X = short.ann$GO_term_IDs, FUN = function(x) {sum(unlist(gregexpr(pattern = "GO:", text = x)) > 0)}, USE.NAMES = F)
	go.num[is.na(short.ann$GO_term_IDs)] <- 0
	
	# calculate data for histograms
	spec.hist <- hist(go.num, breaks = seq(from = go.hist$bins[1]-0.5, to = go.hist$bins[nrow(go.hist)]+0.5), xlim = c(0, 30), main = one.species)
	
	# save to prepared dataframe
	go.hist[, one.species] <- spec.hist$density

	# save statistics to prepared dataframe
	go.stat[1, "legend"] <- "Mean count of GO terms per gene"
	go.stat[1, one.species] <- mean(go.num)
	
	go.stat[2, "legend"] <- "Median count of GO terms per gene"
	go.stat[2, one.species] <- median(go.num)
	
	go.stat[3, "legend"] <- "Percentage of genes without GO term"
	go.stat[3, one.species] <- 100*sum(is.na(short.ann$GO_term_IDs))/nrow(short.ann)
	
	go.stat[4, "legend"] <- "Genes with some GO term"
	# go.stat[4, one.species] <- 100 -100*sum(is.na(short.ann$GO_term_IDs))/nrow(short.ann)
	go.stat[4, one.species] <- paste0(formatC(x = sum(!is.na(short.ann$GO_term_IDs)), big.mark = " "), " (", round(sum(!is.na(short.ann$GO_term_IDs))/nrow(short.ann), 2)*100, "%)")
	
	print(paste(one.species, "done."))
}
## End of loop

## export table with statistics
dir.create(path = "figs_and_stats", recursive = T, showWarnings = F)
write.table(x = go.stat, 
						file = "figs_and_stats/functional_annotation_stats.tsv",
						sep = "\t", row.names = F, na = "", quote = F)

## export table with statistics for publication
go.stat.pub <- go.stat[, c("legend", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
# go.stat.pub[ ,-1] <- round(go.stat.pub[ ,-1], 1)
write.table(x = go.stat.pub, 
						file = "figs_and_stats/functional_annotation_stats_for_pub.tsv",
						sep = "\t", row.names = F, na = "", quote = F)

## Plotting histograms
pdf ("figs_and_stats/n_GO_terms_histogram.pdf", width=7, height=7, onefile = T)

# all species
g.data <- go.hist
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Focus species
colnames(go.hist)
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Arabidopsis_arenosa", "Arabidopsis_lyrata_NCBI", "Arabidopsis_lyrata_Rawat", "Cardamine_amara", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Our annotations
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Resource paper
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, lwd = 2, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lwd = 2, lty = 1)

dev.off()
```

### Notes to the results

As in other statistics, *Cochlearia* seems to have over-predicted genes. Almost 30 % of it's genes doesn't have any GO term annotation, which is much higher than in other species. Most of them are probably not real protein-coding genes.




