---
title: "Functional annotations"
author: "Milos Duchoslav"
date: "2025-11-10"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Length of proteins

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/
mkdir -p protein_length_primary_transcripts

for species_file in primary_transcripts/*.fasta
do
echo "Calculating length for: $species_file"
species=$(sed -E 's-(^.*/)|(\.fasta)--g' <(echo $species_file))
# echo $species

## Calculate length of fasta sequences
# Script from AI, tested
awk '
BEGIN { name=""; seq="" }
/^>/ {
    if (name != "") {
        seq_clean = seq
        sub(/[ *]+$/, "", seq_clean)   # remove trailing * only
        print name "\t" length(seq_clean)
    }
    # extract first word of header without ">"
    header = substr($0, 2)
    split(header, a, /[ \t]/)
    name = a[1]
    seq = ""
    next
}
{
    line = $0
    sub(/[ \t\r\n]+$/, "", line)
    seq = seq line
}
END {
    if (name != "") {
        seq_clean = seq
        sub(/[ *]+$/, "", seq_clean)
        print name "\t" length(seq_clean)
    }
}
' $species_file > protein_length_primary_transcripts/$species"_protein_length.tsv"

done

```


# InterProScan - direct annotation of the genes of desired species (not A. thaliana orthologues)

Inspiration from:

Lian, Qichao, Bruno Huettel, Birgit Walkemeier, Baptiste Mayjonade, Céline Lopez-Roques, Lisa Gil, Fabrice Roux, Korbinian Schneeberger, and Raphael Mercier. “A Pan-Genome of 69 Arabidopsis Thaliana Accessions Reveals a Conserved Genome Structure throughout the Global Species Range.” Nature Genetics 56, no. 5 (May 2024): 982–91. https://doi.org/10.1038/s41588-024-01715-9.:

The resulting gene models were further annotated functionally using InterProScan v5.59-91.0 (ref. 109) (parameters: -f TSV -t p -iprlookup -goterms -pa).

## Interproscan download

https://interproscan-docs.readthedocs.io/en/latest/HowToDownload.html

```{sh}
cd /storage/brno12-cerit/home/duchmil/SW/InterProScan/

wget https://ftp.ebi.ac.uk/pub/software/unix/iprscan/5/5.76-107.0/interproscan-5.76-107.0-64-bit.tar.gz
wget https://ftp.ebi.ac.uk/pub/software/unix/iprscan/5/5.76-107.0/interproscan-5.76-107.0-64-bit.tar.gz.md5

# Recommended checksum to confirm the download was successful:
md5sum -c interproscan-5.76-107.0-64-bit.tar.gz.md5
# Must return *interproscan-5.76-107.0-64-bit.tar.gz: OK*
# If not - try downloading the file again as it may be a corrupted copy.

tar -pxvzf interproscan-5.76-107.0-64-bit.tar.gz
# where:
#     p = preserve the file permissions
#     x = extract files from an archive
#     v = verbosely list the files processed
#     z = filter the archive through gzip
#     f = use archive file

cd interproscan-5.76-107.0/

#Index hmm models
python3 setup.py -f interproscan.properties

# check
module load openjdk
module list
# openjdk/17.0.0_35-intel-19.0.4-udmzouu
./interproscan.sh

# tests
./interproscan.sh -i test_all_appl.fasta -f tsv -dp
./interproscan.sh -i test_all_appl.fasta -f tsv
```

## InterProScan - preparation of input files

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/
mkdir interproscan

cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

## Removing asterisks from protein sequences

# Interproscan doesn't like * in protein sequences, even in the end. We have to remove them from theprotein sequences.

# checking for internal stop codons (* in the middle of sequence)
for file in ../primary_transcripts/*.fasta
do
STOPS=$(awk '/^>/ { if(NR>1) print "";  printf("%s\n",$0); next; } { printf("%s",$0);}  END {printf("\n");}' < $file | grep \*[[:alpha:]] | wc -l)
echo $file $STOPS
done

# checking for * anywhere
grep '\*' ../primary_transcripts/Arabidopsis_arenosa.fasta | wc -l # 33901

# check of sed command
sed -E 's/\*//g' ../primary_transcripts/Noccaea_praecox.fasta | less

## remove all asterisks from protein sequences
# Note: This will remove also internal stop codons in bad proteins.

mkdir -p primary_transcripts_without_stops

for file in ../primary_transcripts/*.fasta
do
  # get filename without path
  filename=$(echo $file | sed 's,^.*/,,')
  
  # remove all asterisks from lines that do not start with >
  sed -E 's/\*//g' ../primary_transcripts/$filename > primary_transcripts_without_stops/$filename
done


```

## InterProScan run

```{sh}
### Metacentrum script

#PBS -N InterProScan_all_species
#PBS -l select=1:ncpus=8:mem=96gb:scratch_local=200gb
#PBS -l walltime=24:00:00 
#PBS -m ae

## It is needed to set the $species variable during submitting the job like this:
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Cardamine_glauca Noccaea_praecox
# do
# echo "Submitting job for species: $species"
# qsub  -v "species=$species" InterProScan_all_species.bash
# done

echo "Species for this job: $species"

# define variables
query_fasta="/storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan/primary_transcripts_without_stops/$species.fasta"
output_dir=/storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

# append a line to a file "jobs_info.txt" containing the ID of the job, the hostname of node it is run on and the path to a scratch directory
# this information helps to find a scratch directory in case the job fails and you need to remove the scratch directory manually 
echo "$PBS_JOBID is running on node `hostname -f` in a scratch directory $SCRATCHDIR" | ts '[%Y-%m-%d %H:%M:%S]' >> $PBS_O_WORKDIR/jobs_info.txt

# test if scratch directory is set
# if scratch directory is not set, issue error message and exit
test -n "$SCRATCHDIR" || { echo >&2 "Variable SCRATCHDIR is not set!"; exit 1; }

# move into scratch directory
cd $SCRATCHDIR 

# load Java
module load openjdk
module list

# run InterProScan
/storage/brno12-cerit/home/duchmil/SW/InterProScan/interproscan-5.76-107.0/interproscan.sh --goterms --pathways --enable-tsv-residue-annot --tempdir $SCRATCHDIR --cpu 8 -i $query_fasta --output-file-base $species

echo "Main calculation done." | ts '[%Y-%m-%d %H:%M:%S]'

# compress output files
pigz -v -p 8 $species.*

# move the output to user's DATADIR or exit in case of failure
mkdir -p $output_dir # will make folder if it doesn't exist
cp -v $species.* $output_dir/ || { echo >&2 "Result file(s) copying failed (with a code $?) !!"; exit 4; }

echo "Copying output file to $output_dir done." | ts '[%Y-%m-%d %H:%M:%S]'

# clean the SCRATCH directory
clean_scratch

# Resources: From 2 h (A. thaliana, probably most precomputed results) up to 24 h (Cochlearia excelsa, many proteins). For most around 10 h, 60-90 % CPU, 50-96 GB memory.

```

### Loop for submitting jobs for several species
```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/metacentrum_scripts

# for species in Aethionema_saxatile Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Arabidopsis_thaliana Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_amara Cardamine_glauca Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Erysimum_linariifolium Euclidium_syriacum Eutrema_salsugineum Noccaea_praecox Odontarrhena_muralis Raphanus_sativus
# for species in Aethionema_saxatile
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Cardamine_amara Cardamine_glauca Erysimum_linariifolium Noccaea_praecox Odontarrhena_muralis
for species in Arabidopsis_thaliana Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Euclidium_syriacum Eutrema_salsugineum Raphanus_sativus
do
echo "Submitting job for species: $species"
qsub  -v "species=$species" InterProScan_all_species.bash
done


```

### Interproscan results

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

# view beginning as table
zcat Arabidopsis_arenosa.tsv.gz | head -n 1000 | column -t -s $'\t' |  less -S

zcat Arabidopsis_arenosa.tsv.gz | wc -l # 315161 (older run with older version of Interproscan: 313362)

zcat Arabidopsis_arenosa.tsv.gz | cut -f 1 | sort | uniq | wc -l # 31121 (out of 33901 proteins) (older run with older version of Interproscan: 31111)

# sites
zless -S Arabidopsis_arenosa.tsv.sites.gz

zcat Arabidopsis_arenosa.tsv.sites.gz | cut -f 1 | sort | uniq | wc -l # 11409 (out of 33901 proteins)
```
#### TSV description
https://interproscan-docs.readthedocs.io/en/latest/OutputFormats.html

The TSV format presents the match data in columns as follows:

    Protein accession (e.g. P51587)

    Sequence MD5 digest (e.g. 14086411a2cdf1c4cba63020e1622579)

    Sequence length (e.g. 3418)

    Analysis (e.g. Pfam / PRINTS / Gene3D)

    Signature accession (e.g. PF09103 / G3DSA:2.40.50.140)

    Signature description (e.g. BRCA2 repeat profile)

    Start location

    Stop location

    Score - is the e-value (or score) of the match reported by member database method (e.g. 3.1E-52)

    Status - is the status of the match (T: true)

    Date - is the date of the run

    InterPro annotations - accession (e.g. IPR002093)

    InterPro annotations - description (e.g. BRCA2 repeat)

    GO annotations with their source(s), e.g. GO:0005515(InterPro)|GO:0006302(PANTHER)|GO:0007195(InterPro,PANTHER). This is an optional column; only displayed if the --goterms option is switched on

    Pathways annotations, e.g. REACT_71. This is an optional column; only displayed if the --pathways option is switched on

#### Analysis description

                      TIGRFAM (XX.X) : TIGRFAMs are protein families based on hidden Markov models (HMMs).
                         SFLD (X) : SFLD is a database of protein families based on hidden Markov models (HMMs).
                  SUPERFAMILY (X.XX) : SUPERFAMILY is a database of structural and functional annotations for all proteins and genomes.
                      PANTHER (XX.X) : The PANTHER (Protein ANalysis THrough Evolutionary Relationships) Classification System is a unique resource that classifies genes by their functions, using published scientific experimental evidence and evolutionary relationships to predict function even in the absence of direct experimental evidence.
                       Gene3D (X.X.X) : Structural assignment for whole genes and genomes using the CATH domain structure database.
                        Hamap (XXXX_XX) : High-quality Automated and Manual Annotation of Microbial Proteomes.
                  ProSiteProfiles (XXX_XX) : PROSITE consists of documentation entries describing protein domains, families and functional sites as well as associated patterns and profiles to identify them.
                        Coils (X.X.X) : Prediction of coiled coil regions in proteins.
                        SMART (X.X) : SMART allows the identification and analysis of domain architectures based on hidden Markov models (HMMs).
                          CDD (X.XX) : CDD predicts protein domains and families based on a collection of well-annotated multiple sequence alignment models.
                       PRINTS (XX.X) : A compendium of protein fingerprints - a fingerprint is a group of conserved motifs used to characterise a protein family.
                        PIRSR (XXXX_XX) : PIRSR is a database of protein families based on hidden Markov models (HMMs) and Site Rules.
                  ProSitePatterns (XXXX_XX) : PROSITE consists of documentation entries describing protein domains, families and functional sites as well as associated patterns and profiles to identify them.
                      AntiFam (X.X) : AntiFam is a resource of profile-HMMs designed to identify spurious protein predictions.
                         Pfam (XX.X) : A large collection of protein families, each represented by multiple sequence alignments and hidden Markov models (HMMs).
                   MobiDBLite (X.X) : Prediction of intrinsically disordered regions in proteins.
                        PIRSF (X.XX) : The PIRSF concept is used as a guiding principle to provide comprehensive and non-overlapping clustering of UniProtKB sequences into a hierarchical order to reflect their evolutionary relationships.
  

# Functional annotations of Arabidopsis thaliana orthologs

## Functions

```{r}
# Extracting of gene/protein anotations from several sources

# This script uses list of A. thaliana genes (AGI codes) and extracts various annotations for them.
# There can be several genes collapsed to one line. In such case, the script collapses also the annotations, if they differ (and can write there the ids for which the annotations apply).


setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
old.par<-par(no.readonly = T)


### FUNKCE

## Funkce k vytahování vlastností genů/proteinů z dalších databází
# Vytahuje informace pro skupiny genů (oddělené např. ";"). Pokud je informace pro celou skupinu stejná, je uvedena jen ta. Pokud je pro geny ze skupiny různá, uvede výčet a v závorce pro které geny.
# Funkce je míněna pro použití s funkcí sapply. Vstup do funkce (x) má být jeden textový řetězec (seznam genů uddělených oddělovačem). Pomocí funkce sapply lze použít na vektor takových textových řetězců.
# Parametry:
# x - textový řetězec s identifikátory genů oddělenými oddělovačem (viz "split")
# table - tabulka identifikátorů genů (resp. vektor) s přiřazenými vlastnostmi; pokud zde bude stejný gen vícekrát, nefunguje přiřazování genů k příslušným vlastnostem (gene.details)
# feature - vlastnosti genů (vektor, data.frame nebo matrix ve stejném pořadí jako "table")
# split - oddělovač použitý v "x" pro řetězení identifikátorů genů
# gene.details - TRUE/FALSE - Vyhodit v závorce, pro které geny platí jaká hodnota?
# collapse - znak, který bude vložen mezi hodnoty v případě více různých vlastností (feature) pro danou skupinu genů

## faster version of function

gene.properties <- function(x, table, feature, split = ", ", gene.details = T, collapse = ", ") {
  # rozdělit "x" podle "split" a vytáhnout k tomu vlastnosti (table - seznam identifikátorů; feature - vlastnosti přiřazené k identifikátorům)
  genes <- unlist(strsplit(x, split = split))
  row.numbers <- sapply(X = genes, FUN = grep, x = table, ignore.case = T)
  feature.v <- as.matrix(feature)[unlist(row.numbers), , drop = F]
  if(nrow(feature.v) == 1) {
    output <- feature.v
  } else if (nrow(feature.v) == 0) {
    output <- rep(NA, ncol(as.matrix(feature)))
  } else {
    output <- vector(mode = "character", length = ncol(as.matrix(feature)))
    for(j in 1:ncol(as.matrix(feature))) {
      feature.ind.v <- feature.v[, j]
      # Je "feature" u všech genů stejná?
      feature.identical <- all(feature.ind.v[1] == feature.ind.v)
      if(is.na(feature.identical)) {
        output[j] <- NA
      } else {
        if(feature.identical) {
          # v případě, že jsou všechny "feature" stejné, vyhoď tuto hodnotu
          output[j] <- feature.ind.v[1]
        } else {
          # v opačném případě načti ke každé hodnotě příslušné geny a ty vyhoď v závorce za každou hodnotou
          # (pokud gene.details = T)
          feature.f <- as.factor(feature.ind.v)
          if(gene.details) {
            feature.m <- matrix(nrow = nlevels(feature.f), ncol = 2)
            feature.m[, 1] <- levels(feature.f)
            for(i in 1:nlevels(feature.f)) {
              genes.level <- genes[feature.f == levels(feature.f)[i]]
              feature.m[i, 2] <- paste(genes.level, collapse = ";")
            }
            output[j] <- paste(paste0(feature.m[, 1], " (", feature.m[, 2], ")"), collapse = collapse)
          }
          else {
            # pokud gene.details = F, vyhoď hodnoty bez genů v závorce
            output[j] <- paste(levels(feature.f), collapse = collapse)
          }
        }
      }
    }
  }
  names(output) <- colnames(feature)
  return(output)
}


# Vectorized, grep-like, no-sapply version
gene.properties_vec <- function(ids_vec,
                                table,
                                feature,
                                split = ", ",
                                gene.details = TRUE,
                                collapse = ", ",
                                ignore.case = TRUE,
                                perl = TRUE) {
  # Prepare feature matrix & dims once
  feature <- as.matrix(feature)
  ncols   <- ncol(feature)
  coln    <- colnames(feature)
  if (is.null(coln)) coln <- paste0("V", seq_len(ncols))

  # Split all groups once
  groups <- strsplit(ids_vec, split, fixed = TRUE)

  # Preallocate output matrix
  out <- matrix(NA_character_, nrow = length(groups), ncol = ncols,
                dimnames = list(NULL, coln))

  # Work through each group (plain for-loop is fastest with preallocation)
  for (g in seq_along(groups)) {
    genes <- groups[[g]]

    # Gather matched row indices and the corresponding genes (to keep mapping)
    idx_all   <- integer(0)
    genes_rep <- character(0)

    # regex-like matching per gene (like grep)
    for (p in genes) {
      if (is.na(p) || p == "") next
      hit <- which(grepl(p, table, ignore.case = ignore.case, perl = perl))
      if (length(hit)) {
        idx_all   <- c(idx_all, hit)
        genes_rep <- c(genes_rep, rep(p, length(hit)))
      }
    }

    # No matches → all NA
    if (!length(idx_all)) {
      next
    }

    fv <- feature[idx_all, , drop = FALSE]

    # Single match → return that row
    if (nrow(fv) == 1L) {
      out[g, ] <- as.character(fv[1, , drop = TRUE])
      next
    }

    # Multiple matches → summarize per column
    for (j in seq_len(ncols)) {
      v <- fv[, j]

      # Preserve original NA behavior:
      # feature.identical <- all(v[1] == v); if (is.na(feature.identical)) NA
      feature_identical <- all(v[1] == v)
      if (is.na(feature_identical)) {
        out[g, j] <- NA_character_
      } else if (feature_identical) {
        out[g, j] <- v[1]
      } else if (gene.details) {
        # List genes for each distinct value (order of appearance)
        # Use tapply on the *values* as groups
        # Ensure stable order of unique(v)
        uv <- unique(v)
        parts <- character(length(uv))
        for (k in seq_along(uv)) {
          val <- uv[k]
          which_k <- which(v == val)
          parts[k] <- sprintf("%s (%s)", val, paste(genes_rep[which_k], collapse = ";"))
        }
        out[g, j] <- paste(parts, collapse = collapse)
      } else {
        # Values only
        out[g, j] <- paste(unique(v), collapse = collapse)
      }
    }
  }

  out
}




### Function for printing messages with time stamps

# Arguments:
# ... - Anything that will be pasted after the time stamp.
# sep - Separator of the arguments printed after the time stamp. Defaults to space.

t.print <- function(..., sep = " ") {
    print(paste0("[", format(x = Sys.time(), format = "%Y-%m-%d %H:%M:%S"), "] ", paste(..., sep = sep)))
}
```

### Testing faster function
```{r}
system.time(
	# extraction of gene properties by the function gene.properties
	aa <- sapply(X = genes$ids[1:1000], FUN = gene.properties, table = subcell[, 1], feature = subcell[, 2], USE.NAMES = F)
)

system.time(
	bb <- gene.properties_vec(
  ids_vec  = genes$ids[1:1000],
  table    = subcell[, 1],
  feature  = subcell[, 2],   # can be vector / matrix / data.frame
  split    = ", ",
  gene.details = TRUE,
  collapse = ", ",
  ignore.case = TRUE,         # same as your grep(..., ignore.case=TRUE)
  perl = TRUE                 # fast regex engine
)
)


cbind(aa, bb)
```


## Reading common files
```{r}
# Read statistics from orthologues
stats.ortho <- read.table("supplemented_orthologues/stats/supplementing_orthologues_stats.tsv", sep = "\t", header = T)


## Gene aliases
# File "gene_aliases_20241001.txt.gz"
# Downloaded on 2024-12-09 from
# https://v2.arabidopsis.org/download_files/Subscriber_Data_Releases/TAIR_Data_20240930/gene_aliases_20241001.txt.gz
# (web page https://v2.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FSubscriber_Data_Releases%2FTAIR_Data_20240930)

aliases <- read.table(file = gzfile("functional_annotation/functional_annotation_data/Araport11/gene_aliases_20241001.txt.gz"), header = T, sep = "\t", quote = "", comment.char = "", fill = T)
summary(aliases)
dim(aliases)

head(aliases)
tail(aliases)

# Are the gene IDs unique?
table(aliases$locus_name)[table(aliases$locus_name)>1] # no
aliases[aliases$locus_name == "AT1G01040", ]

## there are some special characters that make problems
# checking converting
aa <- iconv(aliases$full_name, from = "UTF-8", to = "ASCII", sub = "__AAAA__")
aa[grep("__AAAA__", aa)]
aliases$full_name[grep("__AAAA__", aa)]
# remove special characters
aliases$full_name <- iconv(aliases$full_name, from = "UTF-8", to = "ASCII", sub = "")


## collapsing lines for the same genes
aliases.agg <- aggregate(x = aliases[, 2:3], by = list(aliases$locus_name), FUN = paste, collapse = ", ")
colnames(aliases.agg) <- colnames(aliases)
aliases.agg$full_name <- gsub(pattern = "NULL, |, NULL", replacement = "", x = aliases.agg$full_name)

dim(aliases)
dim(aliases.agg)


## Subcellular predictions
# File "Araport11-Subcellular_Predictions_version_2024_03_09.txt"
# Downloaded on 2024-12-09 from https://www.arabidopsis.org/download/list?dir=Genes%2FAraport11_genome_release

subcell <- read.table(file = "functional_annotation/functional_annotation_data/Araport11/Araport11-Subcellular_Predictions_version_2024_03_09.txt", header = F, sep = "\t")
summary(subcell)
colnames(subcell) <- c("gene", "subcellular.prediction")

# Are the gene IDs unique?
table(subcell$gene)[table(subcell$gene)>1] # yes


## Functional descriptions from arabidopsis.org
# File "Araport11_functional_descriptions_20241001.txt.gz"
# Downloaded on 2024-12-09 from
# https://v2.arabidopsis.org/download_files/Subscriber_Data_Releases/TAIR_Data_20240930/Araport11_functional_descriptions_20241001.txt.gz 
# (web page https://v2.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FSubscriber_Data_Releases%2FTAIR_Data_20240930)

fun.desc <- read.table(file = gzfile("functional_annotation/functional_annotation_data/Araport11/Araport11_functional_descriptions_20241001.txt.gz"), header = T, sep = "\t", quote = "", comment.char = "")
summary(fun.desc)
# Missing values are sometimes "NULL" and sometimes missing.
colSums(fun.desc == "")
colSums(fun.desc == "NULL")
# Making all missing values NULL
fun.desc[fun.desc == ""] <- "NULL"

# Are the gene IDs unique?
table(fun.desc$name)[table(fun.desc$name)>1] # yes

## GO annotations from arabidopsis.org
# Downloaded on 2025-11-11 from
# https://www.arabidopsis.org/download/file?path=Public_Data_Releases/TAIR_Data_20240930/ATH_GO_GOSLIM.txt.gz 
# (web page https://www.arabidopsis.org/download/list?dir=Public_Data_Releases%2FTAIR_Data_20240930)
go.ara <- read.table(file = unz(description = "functional_annotation/functional_annotation_data/Araport11/ATH_GO_GOSLIM.txt.gz", filename = "ATH_GO_GOSLIM.txt"), header = F, sep = "\t", quote = "", comment.char = "", skip = 4)
summary(go.ara)
dim(go.ara)
colnames(go.ara) <- c("locus", "acc_tair", "object_name", "rel_type", "GO_term", "GO_ID", "keyword_id_tair", 
											"aspect", "GOslim_term", "evid_code", "evid_descr", "evid_with", "ref", "annotator", "date")

# Column headers: explanation
# 1. locus name: standard AGI convention name. 
# 2. TAIR accession: the unique identifier for an object in the TAIR database -  the object type is the prefix, followed by a unique accession number(e.g. gene:12345). 
# 3. object name: the name of the object (gene, protein, locus) being annotated.
# 4. relationship type: the relationship between the annotated object and the GO term.
# 5. GO term: the actual string of letters corresponding to the GO ID.
# 6. GO ID: the unique identifier for a GO term. 
# 7. TAIR Keyword ID: the unique identifier for a keyword in the TAIR database.
# 8. Aspect: F=molecular function, C=cellular component, P=biological 13process.
# 9. GOslim term: high level GO term helps in functional categorization.
# 10. Evidence code: three letter code for evidence types (see: http://geneontology.org/page/guide-go-evidence-codes).
# 11. Evidence description: the analysis that was done to support the annotation.
# 12. Evidence with: supporting evidence for IGI, IPI, IC, IEA and ISS annotations.
# 13. Reference: Either a TAIR accession for a reference (reference table: reference_id) or reference from PubMed (e.g. PMID:1234). 
# 14. Annotator: TAIR, TIGR or a TAIR community member.
# 15. Date annotated: date the annotation was made.

go.ara$locus
# Are the gene IDs unique?
table(go.ara$locus)[table(go.ara$locus)>1][1:60]
# collapsing lines for the same genes, choosing just GO_ID
go.ara.agg <- aggregate(x = go.ara[, c("locus", "GO_ID")], by = list(go.ara$locus), FUN = function(x) {paste(unique(x), collapse = ";")})[, -1]
dim(go.ara.agg)


## Metabolic pathways
# File "aracyc_pathways.20230103"
# Downloaded on 2024-12-09 from
# https://plantcyc-ftp.storage.googleapis.com/pmn/Pathways/Data_dumps/PMN15.5_January2023/pathways/aracyc_pathways.20230103
# (web page https://plantcyc.org/)
plantcyc <- read.table("functional_annotation/functional_annotation_data/PlantCyc/aracyc_pathways.20230103", header = T, sep = "\t", quote = "", comment.char = "")
colnames(plantcyc)
head(plantcyc)
# Are the gene IDs unique?
table(plantcyc$Gene.id) # no

## collapsing lines for the same genes
plantcyc.agg <- aggregate(x = plantcyc, by = list(plantcyc$Gene.id), FUN = function(x) {paste(unique(x), collapse = ", ")})[, -1]


## Annotations from Uniprot
# File "uniprotkb_proteome_up000006548_2025_11_11.tsv.gz"
# Downloaded on 2025-11-11 from https://www.uniprot.org/uniprot/?query=proteome%3Aup000006548.
# I selected the desired columns for view and then downloaded that as a table.

uniprot.1 <- read.table(gzfile("functional_annotation/functional_annotation_data/Uniprot/uniprotkb_proteome_up000006548_2025_11_11.tsv.gz"), header = T, sep = "\t", quote = "", comment.char = "")
summary(uniprot.1)
head(uniprot.1)

# Some proteins are annotated by several gene IDs. However, that should not be problem as the gene IDs are searched by grep.
uniprot.1$Gene.Names..ordered.locus.[nchar(uniprot.1$Gene.Names..ordered.locus.) > 9][1:100]



```




## Big loop

```{r}
# read species names
interpro.files <- list.files(path = "interproscan/", pattern = ".tsv.gz")
species <- sub(pattern = ".tsv.gz", replacement = "", x = interpro.files, fixed = T)
# species <- species[-1]
# species <- species[c(1:4, 12:13, 17, 20:21, 6:11, 14:16, 18:19, 22)]
# species <- species[-c(1:11)]

# chosen species for testing of the loop
one.species <- "Aethionema_saxatile"




### Start of the big loop

for(one.species in species) {
	
	  # Save console output to log file
	dir.create(path = "functional_annotation/logs", recursive = T, showWarnings = F)
	log.file <- file(paste0("functional_annotation/logs/", one.species, "_functional_annotation.log"), open = "wt")
  sink(file = log.file, split = T)
  sink(file = log.file, type = "message")
  
  t.print("Starting processing species", one.species)
  
  ## List of genes (AGI codes; there can be multiple genes per line separated by e.g. ";")
  
  # data (gene numbers)
  genes.pre0 <- read.table(file = paste0("supplemented_orthologues/", one.species, "__v__Arabidopsis_thaliana_supplemented.tsv"), sep = "\t", header = T)
  
  genes <- genes.pre0[, 1:2]
  genes$ids <- genes$Arabidopsis_thaliana
  head(genes)
  
  ## 1a. Short version of gene IDs
  short.ids.pre1 <- gsub(pattern = "\\.\\d", replacement = "", x = genes$ids)
  short.ids.pre2 <- strsplit(short.ids.pre1, split = ", ")
  short.ids.l <- lapply(X = short.ids.pre2, FUN = unique)
  short.ids <- sapply(X = short.ids.l, FUN = paste, collapse = ", ")
  short.ids[short.ids == "NA"] <- NA
  
  genes$short.ids <- short.ids
  
  
  ## Gene aliases
  t.print("Extracting gene aliases for", one.species)
  
  # extraction of gene properties by the function gene.properties
  extr.aliases <- t(sapply(X = short.ids, FUN = gene.properties, table = aliases.agg$locus_name, 
                           feature = aliases.agg[, 2:3], gene.details = T, USE.NAMES = F))
  
  
  ## Subcellular predictions
  t.print("Extracting subcellular predictions for", one.species)
  
  # extraction of gene properties by the function gene.properties
  extr.subcell <- sapply(X = genes$ids, FUN = gene.properties, table = subcell[, 1], feature = subcell[, 2], USE.NAMES = F)
  
  
  ## Functional descriptions from arabidopsis.org
    t.print("Extracting functional descriptions from arabidopsis.org for", one.species)
    
  # extraction of gene properties by the function gene.properties
  extr.fun.desc <- t(sapply(X = genes$ids, FUN = gene.properties, table = fun.desc$name, feature = fun.desc[, 2:5], USE.NAMES = F))
  
  ## GO terms from arabidopsis.org
  t.print("Extracting GO terms from arabidopsis.org for", one.species)
  
  # extraction of gene properties by the function gene.properties
  extr.go.ara <- sapply(X = short.ids, FUN = gene.properties, table = go.ara.agg$locus, feature = go.ara.agg$GO_ID, gene.details = F, collapse = ";", USE.NAMES = F)
  
  # each string split by ;, make the items unique and then again collapse them to one string
    used.col <- sapply(X = extr.go.ara, FUN = function(x) {
      paste(unique(unlist(strsplit(x = x, split = ";"))), collapse = ";")
    })
    extr.go.ara <- used.col
  
  
  ## Metabolic pathways
  t.print("Extracting metabolic pathways for", one.species)
  
  # extracting pathways
  pathway <- sapply(X = short.ids, FUN = gene.properties, 
                    table = plantcyc.agg$Gene.id, feature = plantcyc.agg$Pathway.name, gene.details = T, collapse = ", ", 
                    USE.NAMES = F)
  
  
  ## Annotations from Uniprot
  t.print("Extracting annotations from Uniprot for", one.species)
  
  # extraction of gene properties by the function gene.properties
  # system.time(
    uniprot.1.matched <- t(sapply(X = short.ids, FUN = gene.properties, 
                                  table = uniprot.1$Gene.Names..ordered.locus., 
                                  feature = uniprot.1, gene.details = F, collapse = ";", 
                                  USE.NAMES = F))
  # )
  # It can take an hour, would be good to somehow optimize.
  
  
  ## cleaning the matched Uniprot table
  
  uniprot.1.matched.b <- uniprot.1.matched
  # for each column
  for(i in 1:ncol(uniprot.1.matched.b)) {
    used.col <- uniprot.1.matched.b[, i]
    # remove ; at line starts and ends
    used.col <- sub(pattern = "^;|;$", replacement = "", x = used.col)
    # remove double ;;
    used.col <- sub(pattern = ";;", replacement = ";", x = used.col)
    # save the column back
    uniprot.1.matched.b[, i] <- used.col
  }
  
  ## make the Uniprot GO terms and similar columns unique for each protein of original species
  # (removing duplicates in one cell)
  
  # for each column
  # some columns like "Cofactor" use ; in other way, it is better not to include them
  for(i in c(17:27, 31:ncol(uniprot.1.matched.b))) {
    used.col <- uniprot.1.matched.b[, i]
    # remove ; at line starts and ends
    used.col <- sub(pattern = "^;|;$", replacement = "", x = used.col)
    # each string split by ; (can be surrounded by spaces), make the items unique and then again collapse them to one string
    used.col <- sapply(X = used.col, FUN = function(x) {
      paste(unique(unlist(strsplit(x = x, split = " *; *"))), collapse = ";")
    })
    
    uniprot.1.matched.b[, i] <- used.col
  }
  
  ## remove unnecessary Uniprot columns
  # colnames(uniprot.1.matched.b)
  # columns to remove
  cols.rem <- c("Gene.Names..ORF.", 
                "Organism",
                "Reactome",
                "PlantReactome",
                "AlphaFoldDB",
                "KEGG",
                "TAIR",
                "Araport"
  )
  
  uniprot.1.matched.c <- uniprot.1.matched.b[, !colnames(uniprot.1.matched.b) %in% cols.rem]
  
  
  ## Adding link to TAIR and Thalemine
  
  # examples:
  # https://bar.utoronto.ca/thalemine/gene:AT1G31690
  # http://www.arabidopsis.org/servlets/TairObject?name=AT1G06550&type=locus
  
  thalemine.link <- mapply(FUN = paste0, "https://bar.utoronto.ca/thalemine/gene:", short.ids.l, MoreArgs = list(collapse = "; "))
  # remove in the rows with NA
  thalemine.link[is.na(short.ids)] <- NA
  
  tair.link <- mapply(FUN = paste0, "http://www.arabidopsis.org/servlets/TairObject?name=", short.ids.l, "&type=locus", MoreArgs = list(collapse = "; "))
  tair.link[is.na(short.ids)] <- NA
  
  
  ## Completition of the table of annotations
  
  annotations.1 <- cbind(genes.pre0, extr.aliases, extr.subcell, extr.fun.desc, extr.go.ara, pathway, uniprot.1.matched.c, thalemine.link, tair.link)
  colnames(annotations.1)[colnames(annotations.1) == "extr.subcell"] <- "subcellular.prediction"
  colnames(annotations.1)[colnames(annotations.1) == "extr.go.ara"] <- "GO"
  
  # renaming columns (to reflect source)
  colnames(annotations.1)[4:8] <- paste0("OrthoFinder_", colnames(annotations.1)[4:8])
  colnames(annotations.1)[31:38] <- paste0("TAIR_", colnames(annotations.1)[31:38])
  colnames(annotations.1)[39] <- paste0("PlantCyc_", colnames(annotations.1)[39])
  colnames(annotations.1)[40:71] <- paste0("UniProt_", colnames(annotations.1)[40:71])
  
  
  
  ### InterproScan annotation of the original genes (not A. thaliana orthologues)
  
  t.print("Extracting InterproScan annotation for", one.species)
  
  # read gzipped table
  interpro <- read.table(file = gzfile(paste0("interproscan/", one.species, ".tsv.gz")), header = F, sep = "\t", quote = "", comment.char = "")
  
  colnames(interpro) <- c("protein", "md5", "length", "analysis", "signature_accession", 
                          "signature_description", "start", "stop", "score", "status", "run_date", 
                          "interpro_accession", "interpro_description", "GO_term", "pathways")
  
  # ## checks
  # 
  # # how many genes will be added?
  # intepro.genes.uniq <- unique(interpro[, 1])
  # length(intepro.genes.uniq)
  # intepro.genes.new <- intepro.genes.uniq[! intepro.genes.uniq %in% annotations.1[, one.species]]
  # length(intepro.genes.new)
  # 
  # interpro.new <- interpro[interpro[, 1] %in% intepro.genes.new, ]
  
  # list of tools used by InterProScan
  tools.interpro <- sort(unique(interpro$analysis))
  
  # # How many rows and how many different values are there for each tool?
  # for(i in tools.interpro) {
  #   num.lines <- sum(interpro$analysis == i)
  #   num.annot.prot <- length(unique(interpro$protein[interpro$analysis == i]))
  #   nlev.tool <-  nlevels(as.factor(interpro$signature_accession[interpro$analysis == i]))
  #   print(paste(i, "tool - records:", num.lines, "annotated proteins:", num.annot.prot, "levels:", nlev.tool))
  # }
  # 
  # interpro[interpro$analysis == "AntiFam", ][1:10, ]
  
  ## making the interpro table wide (tools as separate columns)
  # prepare the dataframe
  wide.interpro <- data.frame(protein = unique(interpro[, 1]))
  # loop through the tools
  for(i in tools.interpro) {
    # extract rows for the tool
    extr.interpro <- interpro[interpro$analysis == i, ]
    # paste together description, accession and region
    extr.interpro.v <- paste0(extr.interpro$signature_description, " (", extr.interpro$signature_accession,
                              "|", extr.interpro$start, "-", extr.interpro$stop, ")")
    # aggregate if there are several rows for one protein
    extr.interpro.agg <- aggregate(x = extr.interpro.v, by = list(extr.interpro$protein), FUN = paste, collapse = "; ")
    # extr.interpro.v[1:20]
    # extr.interpro.agg[1:20, ]
    
    # add as a new column
    wide.interpro[match(x = extr.interpro.agg[, 1], table = wide.interpro$protein), 
                  which(tools.interpro == i) + 1] <- extr.interpro.agg[, 2]
    # rename the column
    colnames(wide.interpro)[which(tools.interpro == i) + 1] <- paste0("InterProScan_", i)
    
  }
  
  
  ## add interpro accession and description
  
  # extract rows with information
  extr.interpro <- interpro[interpro$interpro_accession != "-", ]
  # paste together description, accession and region
  extr.interpro.v <- paste0(extr.interpro$interpro_description, " (", extr.interpro$interpro_accession,
                            "|", extr.interpro$start, "-", extr.interpro$stop, ")")
  # extr.interpro.v[1:30]
  # aggregate if there are several rows for one protein
  extr.interpro.agg <- aggregate(x = extr.interpro.v, by = list(extr.interpro$protein), FUN = paste, collapse = "; ")
  # add as a new column
  wide.interpro$InterProScan_interpro <- extr.interpro.agg[, 2][match(x = wide.interpro$protein, table = extr.interpro.agg[, 1])]
  

  ## add GO terms
  
  # extract rows with information
  extr.interpro <- interpro[interpro$GO_term != "-", ]
  # remove origin of GO term in parentheses
  extr.interpro.v <- gsub(pattern = "\\([[:alpha:]]*\\)", replacement = "", x = extr.interpro$GO_term)
  # extr.interpro.v[1:30]
  # aggregate if there are several rows for one protein (don't repeat the same GO terms)
  extr.interpro.agg <- aggregate(x = extr.interpro.v, by = list(extr.interpro$protein), FUN = paste, collapse = "|")
  # each string split by |, make the items unique and then again collapse them to one string
  extr.interpro.agg$uniq <- sapply(X = extr.interpro.agg[, 2], FUN = function(x) {
    paste(unique(unlist(strsplit(x = x, split = "|", fixed = T))), collapse = ";")
  })
  # add as a new column
  wide.interpro$InterProScan_GO_term <- extr.interpro.agg$uniq[match(x = wide.interpro$protein, table = extr.interpro.agg[, 1])]
  
  
  
  ## adding Pathways
  
  # extract rows with information
  extr.interpro <- interpro[interpro$pathways != "-", ]
  extr.interpro.v <- extr.interpro$pathways
  # extr.interpro.v[1:3]
  # aggregate if there are several rows for one protein (don't repeat the same GO terms)
  extr.interpro.agg <- aggregate(x = extr.interpro.v, by = list(extr.interpro$protein), FUN = paste, collapse = "|")
  # extr.interpro.agg[1:3, ]
  # each string split by |, make the items unique and then again collapse them to one string
  extr.interpro.agg$uniq <- sapply(X = extr.interpro.agg[, 2], FUN = function(x) {
    paste(unique(unlist(strsplit(x = x, split = "|", fixed = T))), collapse = ";")
  })
  # extr.interpro.agg[1:3, ]
  # add as a new column
  wide.interpro$InterProScan_pathways <- extr.interpro.agg$uniq[match(x = wide.interpro$protein, table = extr.interpro.agg[, 1])]
  
  
  ## Integrating with annotations from A. thaliana
  
  annotations.2 <- merge(x = annotations.1, y = wide.interpro, by.x = one.species, by.y = "protein", all = T)
  
  # View(annotations.2[1:30, c("UniProt_Gene.Ontology.IDs", "InterProScan_GO_term")])
  
  ## Integrate A. thaliana (TAIR and Uniprot) and InterProScan GO terms (target species)
  
  GO_term_IDs <- paste(annotations.2$TAIR_GO, annotations.2$"UniProt_Gene.Ontology.IDs", annotations.2$"InterProScan_GO_term", sep = ";")
  # GO_term_IDs[1:10]
  
  # remove NAs
  GO_term_IDs.2 <- gsub(pattern = "NA", replacement = "", x = GO_term_IDs)
  # GO_term_IDs.2[1:20]
  
  # remove unnecessary ";"
  GO_term_IDs.3 <- gsub(pattern = "^;*|;*$", replacement = "", x = GO_term_IDs.2)
  # GO_term_IDs.3[1:10]
  
  # each string split by "; ", make the items unique and then again collapse them to one string
  GO_term_IDs.uniq <- sapply(X = GO_term_IDs.3, FUN = function(x) {
    paste(unique(unlist(strsplit(x = x, split = ";", fixed = T))), collapse = ";")
  })
  
  annotations.2$GO_term_IDs <- GO_term_IDs.uniq
  
  # annotations.2$GO_term_IDs[1:10]
  
  
  ## reorder columns
  
  # colnames(annotations.2)
  annotations.3 <- annotations.2[, c(1:3, ncol(annotations.2), 4:(ncol(annotations.2)-1))]
  # colnames(annotations.3)
  # colnames(annotations.3) != "InterProScan_pathways"
  
  
  ### Exporting tables
  
  t.print("Exporting tables for", one.species)
  
  # make folders
  dir.create(path = "functional_annotation/1_only_At_orthologues_and_GO", recursive = T, showWarnings = F)
  dir.create(path = "functional_annotation/2_full_without_InterProScan_pathways_tsv_gz", recursive = T, showWarnings = F)
  dir.create(path = "functional_annotation/3_InterProScan_pathways_tsv_gz", recursive = T, showWarnings = F)
  dir.create(path = "functional_annotation/4_full_rds", recursive = T, showWarnings = F)
  
  ## export full table without pathways (pathways are too big)
  
  # gzipped table
  write.table(x = annotations.3[, colnames(annotations.3) != "InterProScan_pathways"], 
              file = gzfile(paste0("functional_annotation/2_full_without_InterProScan_pathways_tsv_gz/", one.species, "_At_orthologues_and_functional_annotations.tsv.gz")), 
              sep = "\t", row.names = F, na = "", quote = F)
  
  # export pathways
  write.table(x = annotations.3[, c(one.species, "InterProScan_pathways")], 
              file = gzfile(paste0("functional_annotation/3_InterProScan_pathways_tsv_gz/", one.species, "_InterProScan_pathways.tsv.gz")), 
              sep = "\t", row.names = F, na = "", quote = F)
  
  # write RDS
  saveRDS(object = annotations.3, 
          file = paste0("functional_annotation/4_full_rds/", one.species, "_At_orthologues_and_functional_annotations_full.rds"))
  
  
  ## export main columns only
  write.table(x = annotations.3[, c(1:4, 43)], 
              file = paste0("functional_annotation/1_only_At_orthologues_and_GO/", one.species, "_At_orthologues_and_GO.tsv"),
              sep = "\t", row.names = F, na = "", quote = F)
  
  t.print(one.species, "finished.")
  
  # Stop saving output to log file
  sink(type = "message")
  sink()
  close(log.file)
}
### End of the big loop

```

# End



Checks of the numbers
```{r}
annotations.3 <- readRDS(file = "functional_annotation/Alyssum_gmelinii_At_orthologues_and_functional_annotations_full.rds")

dim(annotations.3)
sum(!is.na(annotations.3$Arabidopsis_thaliana))
sum(!is.na(annotations.3$single_Arabidopsis_thaliana))

```


# Statistics

Counting GO terms per gene, plotting the histograms

```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
if(!exists("old.par")) old.par<-par(no.readonly = T)

# read the short version of annotation file
ann.files <- list.files(path = "functional_annotation/1_only_At_orthologues_and_GO/", pattern = ".tsv")
species <- sub(pattern = "_At_orthologues_and_GO.tsv", replacement = "", x = ann.files, fixed = T)

# chosen species for testing of the loop
one.species <- "Arabidopsis_arenosa"


### Start of the  loop

# prepare data frames
go.hist <- data.frame(bins = 0:99)
go.stat <- data.frame()

for(one.species in species) {
	short.ann <- read.table(file = paste0("functional_annotation/1_only_At_orthologues_and_GO/", one.species, "_At_orthologues_and_GO.tsv"), header = T, sep = "\t", na.strings = "",
													comment.char = "", quote = "")

	# number of GO terms per gene
	go.num <- sapply(X = short.ann$GO_term_IDs, FUN = function(x) {sum(unlist(gregexpr(pattern = "GO:", text = x)) > 0)}, USE.NAMES = F)
	go.num[is.na(short.ann$GO_term_IDs)] <- 0
	
	# calculate data for histograms
	spec.hist <- hist(go.num, breaks = seq(from = go.hist$bins[1]-0.5, to = go.hist$bins[nrow(go.hist)]+0.5), xlim = c(0, 30), main = one.species)
	
	# save to prepared dataframe
	go.hist[, one.species] <- spec.hist$density

	# save statistics to prepared dataframe
	go.stat[1, "legend"] <- "Mean count of GO terms per gene"
	go.stat[1, one.species] <- mean(go.num)
	
	go.stat[2, "legend"] <- "Median count of GO terms per gene"
	go.stat[2, one.species] <- median(go.num)
	
	go.stat[3, "legend"] <- "Percentage of genes without GO term"
	go.stat[3, one.species] <- 100*sum(is.na(short.ann$GO_term_IDs))/nrow(short.ann)
}
## End of loop

## export table with statistics
dir.create(path = "figs_and_stats", recursive = T, showWarnings = F)
write.table(x = go.stat, 
						file = "figs_and_stats/functional_annotation_stats.tsv",
						sep = "\t", row.names = F, na = "", quote = F)

## Plotting histograms
pdf ("figs_and_stats/n_GO_terms_histogram.pdf", width=7, height=7, onefile = T)

# all species
g.data <- go.hist
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Focus species
colnames(go.hist)
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Arabidopsis_arenosa", "Arabidopsis_lyrata_NCBI", "Arabidopsis_lyrata_Rawat", "Cardamine_amara", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Our annotations
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Resource paper
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, lwd = 2, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lwd = 2, lty = 1)

dev.off()
```

### Results

As in other statistics, *Cochlearia* seems to have over-predicted genes. Almost 30 % of it's genes don't have any GO term annotation, which is much higner than in other species. Most of them are probably not real protein-coding genes.

## Comparison with older version

```{r}
## older version (2.0)
short.ann <- read.table(file = "../brassicaceae_2/R_analysis/Arabidopsis_arenosa_At_orthologues_and_functional_annotations_shortened.tsv", header = T, sep = "\t", na.strings = "",
												comment.char = "", quote = "")

# number of GO terms
go.num <- sapply(X = short.ann$GO_term_IDs, FUN = function(x) {sum(unlist(gregexpr(pattern = "GO:", text = x)) > 0)}, USE.NAMES = F)
go.num[is.na(short.ann$GO_term_IDs)] <- 0

hist(go.num, breaks = (0:100)-0.5, xlim = c(0, 30))
# mean
mean(go.num)
# median
median(go.num)
# Percent of genes without GO term
100*sum(is.na(short.ann$GO_term_IDs))/nrow(short.ann)



```

In A. arenosa, all of the numbers got better. The number of genes without GO terms decreased from 18 % to 7 %.

