---
title: "Functional annotations"
author: "Milos Duchoslav"
date: "2025-11-10"
output: html_document
editor_options: 
  chunk_output_type: console
---


# InterProScan - direct annotation of the genes of desired species (not A. thaliana orthologues)

Inspiration from:

Lian, Qichao, Bruno Huettel, Birgit Walkemeier, Baptiste Mayjonade, Céline Lopez-Roques, Lisa Gil, Fabrice Roux, Korbinian Schneeberger, and Raphael Mercier. “A Pan-Genome of 69 Arabidopsis Thaliana Accessions Reveals a Conserved Genome Structure throughout the Global Species Range.” Nature Genetics 56, no. 5 (May 2024): 982–91. https://doi.org/10.1038/s41588-024-01715-9.:

The resulting gene models were further annotated functionally using InterProScan v5.59-91.0 (ref. 109) (parameters: -f TSV -t p -iprlookup -goterms -pa).



## InterProScan - preparation of input files

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/
mkdir interproscan

cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

## Removing asterisks from protein sequences

# Interproscan doesn't like * in protein sequences, even in the end. We have to remove them from theprotein sequences.

# checking for internal stop codons (* in the middle of sequence)
for file in ../primary_transcripts/*.fasta
do
STOPS=$(awk '/^>/ { if(NR>1) print "";  printf("%s\n",$0); next; } { printf("%s",$0);}  END {printf("\n");}' < $file | grep \*[[:alpha:]] | wc -l)
echo $file $STOPS
done

# checking for * anywhere
grep '\*' ../primary_transcripts/Arabidopsis_arenosa.fasta | wc -l # 33901

# check of sed command
sed -E 's/\*//g' ../primary_transcripts/Noccaea_praecox.fasta | less

## remove all asterisks from protein sequences
# Note: This will remove also internal stop codons in bad proteins.

mkdir -p primary_transcripts_without_stops

for file in ../primary_transcripts/*.fasta
do
  # get filename without path
  filename=$(echo $file | sed 's,^.*/,,')
  
  # remove all asterisks from lines that do not start with >
  sed -E 's/\*//g' ../primary_transcripts/$filename > primary_transcripts_without_stops/$filename
done


```

## InterProScan run

```{sh}
### Metacentrum script

#PBS -N InterProScan_all_species
#PBS -l select=1:ncpus=8:mem=96gb:scratch_local=200gb
#PBS -l walltime=24:00:00 
#PBS -m ae

## It is needed to set the $species variable during submitting the job like this:
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Cardamine_glauca Noccaea_praecox
# do
# echo "Submitting job for species: $species"
# qsub  -v "species=$species" InterProScan_all_species.bash
# done

echo "Species for this job: $species"

# define variables
query_fasta="/storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan/primary_transcripts_without_stops/$species.fasta"
output_dir=/storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

# append a line to a file "jobs_info.txt" containing the ID of the job, the hostname of node it is run on and the path to a scratch directory
# this information helps to find a scratch directory in case the job fails and you need to remove the scratch directory manually 
echo "$PBS_JOBID is running on node `hostname -f` in a scratch directory $SCRATCHDIR" | ts '[%Y-%m-%d %H:%M:%S]' >> $PBS_O_WORKDIR/jobs_info.txt

# test if scratch directory is set
# if scratch directory is not set, issue error message and exit
test -n "$SCRATCHDIR" || { echo >&2 "Variable SCRATCHDIR is not set!"; exit 1; }

# move into scratch directory
cd $SCRATCHDIR 

# load Java
module load openjdk
module list

# run InterProScan
/storage/brno12-cerit/home/duchmil/SW/InterProScan/interproscan-5.76-107.0/interproscan.sh --goterms --pathways --enable-tsv-residue-annot --tempdir $SCRATCHDIR --cpu 8 -i $query_fasta --output-file-base $species

echo "Main calculation done." | ts '[%Y-%m-%d %H:%M:%S]'

# compress output files
pigz -v -p 8 $species.*

# move the output to user's DATADIR or exit in case of failure
mkdir -p $output_dir # will make folder if it doesn't exist
cp -v $species.* $output_dir/ || { echo >&2 "Result file(s) copying failed (with a code $?) !!"; exit 4; }

echo "Copying output file to $output_dir done." | ts '[%Y-%m-%d %H:%M:%S]'

# clean the SCRATCH directory
clean_scratch

# Resources: From 2 h (A. thaliana, probably most precomputed results) up to 24 h (Cochlearia excelsa, many proteins). For most around 10 h, 60-90 % CPU, 50-96 GB memory.

```

### Loop for submitting jobs for several species
```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/metacentrum_scripts

# for species in Aethionema_saxatile Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Arabidopsis_thaliana Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_amara Cardamine_glauca Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Erysimum_linariifolium Euclidium_syriacum Eutrema_salsugineum Noccaea_praecox Odontarrhena_muralis Raphanus_sativus
# for species in Aethionema_saxatile
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Cardamine_amara Cardamine_glauca Erysimum_linariifolium Noccaea_praecox Odontarrhena_muralis
for species in Arabidopsis_thaliana Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Euclidium_syriacum Eutrema_salsugineum Raphanus_sativus
do
echo "Submitting job for species: $species"
qsub  -v "species=$species" InterProScan_all_species.bash
done


```

### Interproscan results

```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/interproscan

# view beginning as table
zcat Arabidopsis_arenosa.tsv.gz | head -n 1000 | column -t -s $'\t' |  less -S

zcat Arabidopsis_arenosa.tsv.gz | wc -l # 315161 (older run with older version of Interproscan: 313362)

zcat Arabidopsis_arenosa.tsv.gz | cut -f 1 | sort | uniq | wc -l # 31121 (out of 33901 proteins) (older run with older version of Interproscan: 31111)

# sites
zless -S Arabidopsis_arenosa.tsv.sites.gz

zcat Arabidopsis_arenosa.tsv.sites.gz | cut -f 1 | sort | uniq | wc -l # 11409 (out of 33901 proteins)
```
#### TSV description
https://interproscan-docs.readthedocs.io/en/latest/OutputFormats.html

The TSV format presents the match data in columns as follows:

    Protein accession (e.g. P51587)

    Sequence MD5 digest (e.g. 14086411a2cdf1c4cba63020e1622579)

    Sequence length (e.g. 3418)

    Analysis (e.g. Pfam / PRINTS / Gene3D)

    Signature accession (e.g. PF09103 / G3DSA:2.40.50.140)

    Signature description (e.g. BRCA2 repeat profile)

    Start location

    Stop location

    Score - is the e-value (or score) of the match reported by member database method (e.g. 3.1E-52)

    Status - is the status of the match (T: true)

    Date - is the date of the run

    InterPro annotations - accession (e.g. IPR002093)

    InterPro annotations - description (e.g. BRCA2 repeat)

    GO annotations with their source(s), e.g. GO:0005515(InterPro)|GO:0006302(PANTHER)|GO:0007195(InterPro,PANTHER). This is an optional column; only displayed if the --goterms option is switched on

    Pathways annotations, e.g. REACT_71. This is an optional column; only displayed if the --pathways option is switched on

#### Analysis description

                      TIGRFAM (XX.X) : TIGRFAMs are protein families based on hidden Markov models (HMMs).
                         SFLD (X) : SFLD is a database of protein families based on hidden Markov models (HMMs).
                  SUPERFAMILY (X.XX) : SUPERFAMILY is a database of structural and functional annotations for all proteins and genomes.
                      PANTHER (XX.X) : The PANTHER (Protein ANalysis THrough Evolutionary Relationships) Classification System is a unique resource that classifies genes by their functions, using published scientific experimental evidence and evolutionary relationships to predict function even in the absence of direct experimental evidence.
                       Gene3D (X.X.X) : Structural assignment for whole genes and genomes using the CATH domain structure database.
                        Hamap (XXXX_XX) : High-quality Automated and Manual Annotation of Microbial Proteomes.
                  ProSiteProfiles (XXX_XX) : PROSITE consists of documentation entries describing protein domains, families and functional sites as well as associated patterns and profiles to identify them.
                        Coils (X.X.X) : Prediction of coiled coil regions in proteins.
                        SMART (X.X) : SMART allows the identification and analysis of domain architectures based on hidden Markov models (HMMs).
                          CDD (X.XX) : CDD predicts protein domains and families based on a collection of well-annotated multiple sequence alignment models.
                       PRINTS (XX.X) : A compendium of protein fingerprints - a fingerprint is a group of conserved motifs used to characterise a protein family.
                        PIRSR (XXXX_XX) : PIRSR is a database of protein families based on hidden Markov models (HMMs) and Site Rules.
                  ProSitePatterns (XXXX_XX) : PROSITE consists of documentation entries describing protein domains, families and functional sites as well as associated patterns and profiles to identify them.
                      AntiFam (X.X) : AntiFam is a resource of profile-HMMs designed to identify spurious protein predictions.
                         Pfam (XX.X) : A large collection of protein families, each represented by multiple sequence alignments and hidden Markov models (HMMs).
                   MobiDBLite (X.X) : Prediction of intrinsically disordered regions in proteins.
                        PIRSF (X.XX) : The PIRSF concept is used as a guiding principle to provide comprehensive and non-overlapping clustering of UniProtKB sequences into a hierarchical order to reflect their evolutionary relationships.
  

# Functional annotations of Arabidopsis thaliana orthologs


## Compilation of annotations from several sources

List of species
```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
old.par<-par(no.readonly = T)

# read species names
interpro.files <- list.files(path = "interproscan/", pattern = ".tsv.gz")
species <- sub(pattern = ".tsv.gz", replacement = "", x = interpro.files, fixed = T)
species
```

### Testing of script
```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3

# load R
module load r/4.1.3-gcc-10.2.1-6xt26dl

Rscript --verbose "functional_annotation_compilation_executable.r" Aethionema_saxatile
```

### Bash script for running of R script (compilation of functional annotations)
```{sh}
### Metacentrum script

#PBS -N functional_annotation_compilation
#PBS -l select=1:ncpus=1:mem=32gb:scratch_local=200gb
#PBS -l walltime=24:00:00 
#PBS -m ae

## It is needed to set the $species variable during submitting the job like this:
# for species in Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Cardamine_glauca Noccaea_praecox
# do
# echo "Submitting job for species: $species"
# qsub  -v "species=$species" InterProScan_all_species.bash
# done

echo "Species for this job: $species" | ts '[%Y-%m-%d %H:%M:%S]'

# append a line to a file "jobs_info.txt" containing the ID of the job, the hostname of node it is run on and the path to a scratch directory
# this information helps to find a scratch directory in case the job fails and you need to remove the scratch directory manually 
echo "$PBS_JOBID is running on node `hostname -f` in a scratch directory $SCRATCHDIR" | ts '[%Y-%m-%d %H:%M:%S]' >> $PBS_O_WORKDIR/jobs_info.txt

# move to working dir
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3

# load R
module load r/4.1.3-gcc-10.2.1-6xt26dl

# run R script
Rscript --verbose "functional_annotation_compilation_executable.r" $species || { echo >&2 "Rscript failed!"; exit 1; }

echo "Run ended: $species" | ts '[%Y-%m-%d %H:%M:%S]'

# clean the SCRATCH directory
clean_scratch

# Resources: 100 % CPU, up to 2 h, 6 GB memory.
```

### Loop for submitting jobs for several species
```{sh}
cd /storage/brno12-cerit/home/duchmil/Brassicaceae_orthology/brassicaceae_3/metacentrum_scripts

# for species in Aethionema_saxatile Alyssum_gmelinii Arabidopsis_arenosa Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_amara Cardamine_glauca Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Erysimum_linariifolium Euclidium_syriacum Eutrema_salsugineum Noccaea_praecox Odontarrhena_muralis Raphanus_sativus
# for species in Alyssum_gmelinii Arabidopsis_arenosa
for species in Arabidopsis_lyrata_NCBI Arabidopsis_lyrata_Rawat Arabis_alpina Brassica_napus Brassica_oleracea Brassica_rapa Camelina_sativa Capsella_rubella Cardamine_amara Cardamine_glauca Cardamine_hirsuta Cochlearia_excelsa Conringia_planisiliqua Erysimum_linariifolium Euclidium_syriacum Eutrema_salsugineum Noccaea_praecox Odontarrhena_muralis Raphanus_sativus
do
echo "Submitting job for species: $species"
qsub  -v "species=$species" functional_annotation_compilation.bash
done


```

# Adding reliability table for species annotated by me

```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
if(!exists("old.par")) old.par<-par(no.readonly = T)

# dirs with annotations
dirs.list <- list.dirs(path = "D:/!ecolgen/annotations/", recursive = F)
# support tables
files.list <- list.files(path = paste0(dirs.list, "/final_files/"), pattern = "_protein_coding_genes_support.tsv", full.names = T)

## preparation of dataframe for histogram data
# set boundaries and coarsness of frequency calculation
l.min <- 0
l.max <- 10000
l.by <- 25
# prepare data frames
length.hist <- data.frame(bins = seq(from = l.min + (l.by / 2), to = l.max - (l.by / 2), by = l.by))
length.hist.1 <- data.frame(bins = seq(from = l.min + (l.by / 2), to = l.max - (l.by / 2), by = l.by))
length.hist.2 <- data.frame(bins = seq(from = l.min + (l.by / 2), to = l.max - (l.by / 2), by = l.by))

# write files?
write.files <- F

for(i in seq_along(files.list)) {
	
	supp <- read.table(file = files.list[i], header = T, sep = "\t")
	# supp <- read.table(file = "final_files/Odontarrhena_muralis_CUNI_V1_annotation_v1.0_protein_coding_genes_support.tsv", header = T, sep = "\t")
	
	summary(supp)
	
	colnames(supp) <- paste0("support_", colnames(supp))
	
	# Extract species from file name
	file.name <- gsub(pattern = "^.*/", replacement = "", x = files.list[i])
	one.species <- gsub(pattern = "(^[^_]*_[^_]*).*", replacement = "\\1", x = file.name)
	
	# read RDS
	annotations.3 <- readRDS(file = paste0("functional_annotation/4_full_rds/", one.species, "_At_orthologues_and_functional_annotations_full.rds"))
	
	colnames(annotations.3)
	
	# merge tables
	annotations.4 <- merge(x = annotations.3, y = supp, by.x = one.species, by.y = "support_Gene", all = T)
	
	# column after which to move the supp columns
	d <- 2
	# move supp columns
	annotations.5 <- annotations.4[, c(1:d, (ncol(annotations.3) + 1):ncol(annotations.4), (d + 1):ncol(annotations.3))]
	
	colnames(annotations.5)
	
	# save merged table
	if(write.files) {
		# make folder
		dir.create(path = "functional_annotation/5_full_with_support_for_my_annotations_rds", recursive = T, showWarnings = F)
		# write RDS
		saveRDS(object = annotations.5, 
						file = paste0("functional_annotation/5_full_with_support_for_my_annotations_rds/", one.species, "_support_and_At_orthologues_and_functional_annotations.rds"))
	}
	
	# Genes supported by proteins from at least one species
	support_any_protein <- rowSums(annotations.5[, c("support_A.lyrata.pep", "support_A.thaliana.pep", "support_B.rapa.pep")]) >= 1

	# Genes with any support (either protein from at least one species or transcripts assembled from RNAseq)
	support_any <- rowSums(annotations.5[, c("support_Assembled_transcripts", "support_A.lyrata.pep", "support_A.thaliana.pep", "support_B.rapa.pep")]) >= 1

	# Genes with support of proteins from all species and transcripts
	support_all <- rowSums(annotations.5[, c("support_Assembled_transcripts", "support_A.lyrata.pep", "support_A.thaliana.pep", "support_B.rapa.pep")]) == 4
	
	# calculate data for histograms
	# hist(annotations.5$prot_length[annotations.5$support_Assembled_transcripts == 1], breaks = 50)
	# hist(annotations.5$prot_length[support_any == 1], breaks = 50)
	
	# all proteins
	spec.hist <- hist(annotations.5$prot_length, breaks = seq(from = l.min, to = l.max, by = l.by), xlim = c(0, 3000), main = one.species)
	# proteins with any support (either protein from at least one species or transcripts assembled from RNAseq)
	spec.hist.1 <- hist(annotations.5$prot_length[support_any == 1], breaks = seq(from = l.min, to = l.max, by = l.by), xlim = c(0, 3000), main = one.species)
	# proteins with support of proteins from all species and transcripts
	spec.hist.2 <- hist(annotations.5$prot_length[support_all == 1], breaks = seq(from = l.min, to = l.max, by = l.by), xlim = c(0, 3000), main = one.species)
	
	# save to prepared dataframe
	length.hist[, one.species] <- spec.hist$density
	length.hist.1[, one.species] <- spec.hist.1$density
	length.hist.2[, one.species] <- spec.hist.2$density
	
}

## Plotting histograms
pdf ("figs_and_stats/prot_length_histogram_supported_proteins.pdf", width=7, height=7, onefile = T)

# plots input
g.input <- c("length.hist", "length.hist.1", "length.hist.2")
g.sub <- c("All genes", 
						"Genes supported by aligned protein from at least one species or by transcript",
						"Genes supported by aligned proteins from all three species and by transcript")
# all species
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19, lwd = 2,
					main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length (amino acid residues)")
	legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1, lwd = 2)
	mtext(text = g.sub[i])
}

# Resource paper
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = rainbow(ncol(g.data)-1), pch = 19,
				main = "Protein length spectrum", ylab = "Frequency", xlab = "Protein length (amino acid residues)", lwd = 2)
	legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1, lwd = 2)
	mtext(text = g.sub[i])
}
dev.off()

## Plot for publication
# use palette from ColorBrewer
require(RColorBrewer)
display.brewer.all(n = 5, type = "qual", colorblindFriendly = T)

# plots input
g.input <- c("length.hist", "length.hist.1", "length.hist.2")
g.sub <- c("All genes", 
						"Genes supported by aligned protein from at least one species or by transcript",
						"Genes supported by aligned proteins from all three species and by transcript")

# Resource paper
pdf ("figs_and_stats/prot_length_histogram_for_publication_supported_proteins.pdf", width=7, height=7, onefile = T)
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3))
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	mtext(text = g.sub[i], line = 2)
}
dev.off()

pdf ("figs_and_stats/prot_length_histogram_for_publication_supported_proteins_2.pdf", width=10, height=5, onefile = T)
par(mfrow = c(1, 2))
for(i in 1:2) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3), las = 1)
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	# mtext(text = g.sub[i], line = 2)
	mtext(text = c("A", "B")[i], line = 2, adj = 0, font = 2, cex = 2)
}
par(old.par)
dev.off()

svg ("figs_and_stats/prot_length_histogram_for_publication_supported_proteins_3.svg", width=180/25.4, height=90/25.4, onefile = T)
par(mfrow = c(1, 2), cex = 0.75)
for(i in 1:2) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 2000), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3), las = 1)
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	# mtext(text = g.sub[i], line = 2)
		# mtext(text = c("a", "b")[i], line = 2, adj = 0, font = 2, cex = 2)
	mtext(text = c(expression(bold("a") * " All gene models"), expression(bold("b") * " Gene models with external evidence"))[i], line = 1.5, adj = 0, font = 1, cex = 1, outer = F)
}
par(old.par)
dev.off()

# just range of short proteins
for(i in seq_along(g.input)) {
	g.data <- get(g.input[i])[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
	# Change to %
	g.data[, -1] <- g.data[, -1] * 100
	matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 500), lty = 1, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), pch = 19,
					main = "", ylab = "Frequency (%)", xlab = "Protein length (amino acid residues)", lwd = 2, ylim = c(0, 0.3))
	g.legend <- gsub(pattern = "_", replacement = " ", x = colnames(g.data[-1]))
	legend(x = "topright", legend = g.legend, col = brewer.pal(n = ncol(g.data)-1, name = "Dark2"), lwd = 2, lty = 1, text.font = 3)
	mtext(text = g.sub[i], line = 2)
}

```



# Statistics of functional annotation

## GO terms

Counting GO terms per gene, plotting the histograms

```{r}
setwd("D:/!ecolgen/Brassicaceae_orthology/brassicaceae_3/")
if(!exists("old.par")) old.par<-par(no.readonly = T)

# read the short version of annotation file
ann.files <- list.files(path = "functional_annotation/1_only_At_orthologues_and_GO/", pattern = ".tsv")
species <- sub(pattern = "_At_orthologues_and_GO.tsv", replacement = "", x = ann.files, fixed = T)

# chosen species for testing of the loop
one.species <- "Arabidopsis_arenosa"


### Start of the  loop

# prepare data frames
go.hist <- data.frame(bins = 0:99)
go.stat <- data.frame()

for(one.species in species) {
	short.ann <- read.table(file = paste0("functional_annotation/1_only_At_orthologues_and_GO/", one.species, "_At_orthologues_and_GO.tsv"), header = T, sep = "\t", na.strings = "",
													comment.char = "", quote = "")

	# number of GO terms per gene
	go.num <- sapply(X = short.ann$GO_term_IDs, FUN = function(x) {sum(unlist(gregexpr(pattern = "GO:", text = x)) > 0)}, USE.NAMES = F)
	go.num[is.na(short.ann$GO_term_IDs)] <- 0
	
	# calculate data for histograms
	spec.hist <- hist(go.num, breaks = seq(from = go.hist$bins[1]-0.5, to = go.hist$bins[nrow(go.hist)]+0.5), xlim = c(0, 30), main = one.species)
	
	# save to prepared dataframe
	go.hist[, one.species] <- spec.hist$density

	# save statistics to prepared dataframe
	go.stat[1, "legend"] <- "Mean count of GO terms per gene"
	go.stat[1, one.species] <- mean(go.num)
	
	go.stat[2, "legend"] <- "Median count of GO terms per gene"
	go.stat[2, one.species] <- median(go.num)
	
	go.stat[3, "legend"] <- "Percentage of genes without GO term"
	go.stat[3, one.species] <- 100*sum(is.na(short.ann$GO_term_IDs))/nrow(short.ann)
	
	go.stat[4, "legend"] <- "Genes with some GO term"
	# go.stat[4, one.species] <- 100 -100*sum(is.na(short.ann$GO_term_IDs))/nrow(short.ann)
	go.stat[4, one.species] <- paste0(formatC(x = sum(!is.na(short.ann$GO_term_IDs)), big.mark = " "), " (", round(sum(!is.na(short.ann$GO_term_IDs))/nrow(short.ann), 2)*100, "%)")
	
	print(paste(one.species, "done."))
}
## End of loop

## export table with statistics
dir.create(path = "figs_and_stats", recursive = T, showWarnings = F)
write.table(x = go.stat, 
						file = "figs_and_stats/functional_annotation_stats.tsv",
						sep = "\t", row.names = F, na = "", quote = F)

## export table with statistics for publication
go.stat.pub <- go.stat[, c("legend", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
# go.stat.pub[ ,-1] <- round(go.stat.pub[ ,-1], 1)
write.table(x = go.stat.pub, 
						file = "figs_and_stats/functional_annotation_stats_for_pub.tsv",
						sep = "\t", row.names = F, na = "", quote = F)

## Plotting histograms
pdf ("figs_and_stats/n_GO_terms_histogram.pdf", width=7, height=7, onefile = T)

# all species
g.data <- go.hist
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Focus species
colnames(go.hist)
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Arabidopsis_arenosa", "Arabidopsis_lyrata_NCBI", "Arabidopsis_lyrata_Rawat", "Cardamine_amara", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Our annotations
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Alyssum_gmelinii", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lty = 1)

# Resource paper
g.data <- go.hist[, c("bins", "Aethionema_saxatile", "Cardamine_glauca", "Erysimum_linariifolium", "Noccaea_praecox", "Odontarrhena_muralis")]
matplot(x = g.data$bins, y = g.data[, -1], type = "l", xlim = c(0, 30), lty = 1, lwd = 2, col = rainbow(ncol(g.data)-1),
				main = "Number of GO terms per gene", ylab = "Frequency", xlab = "Number of GO terms per gene")
legend(x = "topright", legend = colnames(g.data[-1]), col = rainbow(ncol(g.data)-1), lwd = 2, lty = 1)

dev.off()
```

### Results

As in other statistics, *Cochlearia* seems to have over-predicted genes. Almost 30 % of it's genes doesn't have any GO term annotation, which is much higher than in other species. Most of them are probably not real protein-coding genes.


